### 2025-01-10 MAT185 Lecture Notes/main.tex

\documentclass[11pt]{article}
\input{preamble}
\begin{document}
\thispagestyle{empty}
{\LARGE \bf MAT 185 Lecture Notes}\\
{\large Hei Shing Cheung}\\
Linear Algebra, Winter 2024 \hfill MAT 185\\
\section{Vector Space}
\begin{definition}[Vector Space in $\mathbb{R}$]
    A vectors space,${V}$, over $\mathbb{R}$ is a collection of \textbf{object} $\textbf{v} \in V$ s.t. the follow axioms are followed
    \begin{enumerate}
        \item \textbf{Addition Axioms}
        \begin{enumerate}
            \item \textbf{Closure Under Addition:} \\
            $\textbf{x}, \textbf{y} \in V \implies \textbf{x} + \textbf{y} \in V$

            \item \textbf{Associativity of Addition:} \\
            $\textbf{x}, \textbf{y}, \textbf{z} \in V \implies (\textbf{x} + \textbf{y}) + \textbf{z} = \textbf{x} + (\textbf{y} + \textbf{z})$

            \item \textbf{Existence of Additive Identity:} \\
            $\exists \, \textbf{0} \in V$ such that $\textbf{x} \in V \implies \textbf{x} + \textbf{0} = \textbf{x}$

            \item \textbf{Existence of Additive Inverse:} \\
            $\forall \, \textbf{x} \in V, \, \exists \, \textbf{-x} \in V$ such that $\textbf{x} + \textbf{-x} = \textbf{0}$
        \end{enumerate}

        \item \textbf{Scalar Multiplication Axioms}
        \begin{enumerate}
            \item \textbf{Closure Under Scalar Multiplication:} \\
            $\forall \, \textbf{x} \in V \, \text{and} \, \forall \, \alpha \in \mathbb{R}, \, \alpha \textbf{x} \in V$

            \item \textbf{Associativity of Scalar Multiplication:} \\
            $\forall \, \textbf{x} \in V \, \text{and} \, \forall \, \alpha, \beta \in \mathbb{R}, \, (\alpha \beta) \textbf{x} = \alpha (\beta \textbf{x})$

            \item \textbf{Distributive Property of Scalar Multiplication:} \\
            $\forall \, \textbf{x} \in V \, \text{and} \, \forall \, \alpha, \beta \in \mathbb{R}, \, (\alpha + \beta) \textbf{x} = \alpha \textbf{x} + \beta \textbf{x}$

            \item \textbf{Existence of Multiplicative Identity:} \\
            $\forall \, \textbf{x} \in V, \, 1 \textbf{x} = \textbf{x}$
        \end{enumerate}
    \end{enumerate}

    \paragraph{Note} It could be shown that the axiom imply the commutativity of in addition, namely $\forall \textbf{x} \implies \textbf{y} \in \mathbb{R}$, $\textbf{x} + \textbf{y} = \textbf{y} + \textbf{x}$
\end{definition}
\paragraph{Example} $\mathbb{R}^n$ is a vector space over $\mathbb{R}$, where $\textbf{x} = (x_1, x_2, \ldots, x_n)$ and $\textbf{y} = (y_1, y_2, \ldots, y_n)$, then $\textbf{x} + \textbf{y} = (x_1 + y_1, x_2 + y_2, \ldots, x_n + y_n)$ and $\alpha\textbf{x} = (\alpha x_1, \alpha x_2, \ldots, \alpha x_n)$
\begin{theorem}[Cancelation, Part 1]
    Let $V$ be a vector space and $\textbf{x}, \textbf{y}, \textbf{z} \in V$, then:
    \begin{align*}
        \textbf{x} + \textbf{z} &= \textbf{y} + \textbf{z} \\
        \textbf{x} &= \textbf{y}
    \end{align*} 
\end{theorem}
\begin{proof}
    \begin{align*}
        \textbf{x} + \textbf{z} &= \textbf{y} + \textbf{z} \\
        \intertext{From additive inverse we know that \textbf{-z} exists}
        (\textbf{x} + \textbf{z}) + \textbf{-z} &= (\textbf{y} + \textbf{z}) + \textbf{-z} \\
        \intertext{By order of addition we have: }
        \textbf{x} + (\textbf{z} + \textbf{-z}) &= \textbf{y} + (\textbf{z} + \textbf{-z}) \\
        \textbf{x} + \textbf{0} &= \textbf{y} + \textbf{0} \\
        \intertext{By additive identity we have: }
        \textbf{x} &= \textbf{y}
    \end{align*}
\end{proof}
\begin{theorem}[Cancelation, Part 2]
    Let $V$ be a vector space and $\textbf{x}, \textbf{y}, \textbf{z} \in V$, then:
    \begin{align*}
        \textbf{z} + \textbf{x} &= \textbf{z} + \textbf{y} \\
        \textbf{x} &= \textbf{y}
    \end{align*} 
\end{theorem}
To proof that, it would require the following propostion:
\begin{lemma}
    Let $V$ be a vector space and $\textbf{z} \in V$, then $\textbf{-z} + \textbf{z} = 0 $
\end{lemma}
\begin{proof}
    \begin{align*}
        \intertext{We know:}
        \textbf{-z} + \textbf{z} &= (\textbf{-z} + \textbf{z}) + 0\\
        &= (\textbf{-z} + \textbf{z}) + (\textbf{-z} + \textbf{-(-z)}) \\
        \intertext{By the order of additon, }
        &= \textbf{-z} + (\textbf{z} + (\textbf{-z} + \textbf{-(-z)})) \\
        \intertext{Again, by the order of additon, }
        &= \textbf{-z} + ((\textbf{z} + \textbf{-z}) + \textbf{-(-z)}) \\
        &= \textbf{-z} + \textbf{-(-z)} \\
        &= 0
    \end{align*}
\end{proof}
Now, to prove the part 2 of the Cancelation Theorem:
\begin{proof}
    \begin{align*}
        \textbf{z} + \textbf{x} &= \textbf{z} + \textbf{y} \\
        \intertext{From additive inverse we know that \textbf{-z} exists}
        \textbf{-z} + (\textbf{z} + \textbf{x}) &= \textbf{-z} + (\textbf{z} + \textbf{y})\\
        (\textbf{-z} + \textbf{z}) + \textbf{x} &= (\textbf{-z} + \textbf{z}) + \textbf{y}\\
        \intertext{From above, we have $\textbf{-z} + \textbf{z} = 0$}
        0 + \textbf{x} &= 0 + \textbf{y} \\
        \textbf{x} &= \textbf{y}
    \end{align*}
\end{proof}
\begin{lemma}[Inverse of an inverse]
    Let $V$ be a vector space and $\textbf{x} \in V$, then:
    \begin{align*}
        \textbf{-(-x)} &= \textbf{x}
    \end{align*}
\end{lemma}
\begin{proof}
    Assume $0$, $0^\star$ are the additive identity of $V$ and $\textbf{-x}$, $\textbf{-x}^\star$ are the additive inverse of $\textbf{x}$. We have:
    \begin{equation*}
        u+0 = u + 0^\star
    \end{equation*}
    By Cancelation Theorem, we have $0 = 0^\star$. Thus, the additive identity is unique. Now, we have:
    \begin{align*}
        \textbf{x} + \textbf{-x} &= 0 \\
        \textbf{x} + \textbf{-x}^\star &= 0
        \intertext{Thus, by the uniqueness of the additive inverse, we have:}
        \textbf{x} + \textbf{-x} &= \textbf{x} + \textbf{-x}^\star
        \intertext{By the Cancelation Theorem, we have:}
        \textbf{-x} &= \textbf{-x}^\star
    \end{align*}
    Now, we have that the additive inverse is unique. Thus, $\textbf{-(-x)}$ must be unique and nessarily $\textbf{x}$. 
\end{proof}
\paragraph{Results from above}
\begin{enumerate}
    \item The additive identity is unique
    \item The additive inverse is unique
\end{enumerate}
\begin{definition}[Subtraction]
    Let $V$ be a vector space and $\textbf{x}, \textbf{y} \in V$, then:
    \begin{align*}
        \textbf{x} - \textbf{y} &\defeq \textbf{x} + \textbf{-y}
    \end{align*}
\end{definition}
\begin{theorem}[Commutativity of Addition]
    Let $V$ be a vector space and $\textbf{x}, \textbf{y} \in V$, then:
    \begin{align*}
        \textbf{x} + \textbf{y} &= \textbf{y} + \textbf{x}
    \end{align*} 
\end{theorem}
\begin{proof}
    \begin{align*}
        \textbf{x} + \textbf{y} &=  
    \end{align*}
\end{proof}   
\end{document}