### 2025-01-16 ESC195 Lecture Notes/AppendixItems/trigDeri.tex

\section{Trigonometric Derivatives}
$$
\setlength{\tabcolsep}{8pt} % Adjust column spacing
\renewcommand{\arraystretch}{1.5} % Adjust row spacing
\begin{array}{|l|l|}
\hline
\textbf{Trigonometric Function} & \textbf{Inverse Trigonometric Function} \\ \hline
\dfrac{d}{dx} \text{sin}(x) = \text{cos}(x) & \dfrac{d}{dx} \text{arcsin}(x) = \dfrac{1}{\sqrt{1 - x^2}} \\ \hline
\dfrac{d}{dx} \text{cos}(x) = -\text{sin}(x) & \dfrac{d}{dx} \text{arccos}(x) = \dfrac{-1}{\sqrt{1 - x^2}} \\ \hline
\dfrac{d}{dx} \text{tan}(x) = \text{sec}^2(x) & \dfrac{d}{dx} \text{arctan}(x) = \dfrac{1}{1 + x^2} \\ \hline
\dfrac{d}{dx} \text{cot}(x) = -\text{csc}^2(x) & \dfrac{d}{dx} \text{arccot}(x) = \dfrac{-1}{1 + x^2} \\ \hline
\dfrac{d}{dx} \text{sec}(x) = \text{sec}(x)\text{tan}(x) & \dfrac{d}{dx} \text{arcsec}(x) = \dfrac{1}{|x|\sqrt{x^2 - 1}} \\ \hline
\dfrac{d}{dx} \text{csc}(x) = -\text{csc}(x)\text{cot}(x) & \dfrac{d}{dx} \text{arccsc}(x) = \dfrac{-1}{|x|\sqrt{x^2 - 1}} \\ \hline
\hline
\textbf{Hyperbolic Function} & \textbf{Inverse Hyperbolic Function} \\ \hline
\dfrac{d}{dx} \text{sinh}(x) = \text{cosh}(x) & \dfrac{d}{dx} \text{arsinh}(x) = \dfrac{1}{\sqrt{x^2 + 1}} \\ \hline
\dfrac{d}{dx} \text{cosh}(x) = \text{sinh}(x) & \dfrac{d}{dx} \text{arcosh}(x) = \dfrac{1}{\sqrt{x^2 - 1}} \quad (x > 1) \\ \hline
\dfrac{d}{dx} \text{tanh}(x) = \text{sech}^2(x) & \dfrac{d}{dx} \text{artanh}(x) = \dfrac{1}{1 - x^2} \quad (|x| < 1) \\ \hline
\dfrac{d}{dx} \text{coth}(x) = -\text{csch}^2(x) & \dfrac{d}{dx} \text{arcoth}(x) = \dfrac{1}{1 - x^2} \quad (|x| > 1) \\ \hline
\dfrac{d}{dx} \text{sech}(x) = -\text{sech}(x)\text{tanh}(x) & \dfrac{d}{dx} \text{arsech}(x) = \dfrac{-1}{x\sqrt{1 - x^2}} \quad (0 < x \leq 1) \\ \hline
\dfrac{d}{dx} \text{csch}(x) = -\text{csch}(x)\text{coth}(x) & \dfrac{d}{dx} \text{arcsch}(x) = \dfrac{-1}{|x|\sqrt{x^2 + 1}} \\ \hline
\end{array}
$$