### 2025-02-06 ESC195 Lecture Notes/main.tex
feb 6

\documentclass[11pt]{article}
\input{preamble}
\begin{document}
\thispagestyle{empty}
{\LARGE \bf ESC 195 Lecture Notes}\\
{\large Hei Shing Cheung}\\
Caculus II, Winter 2024 \hfill ESC 195\\
\\
The up-to-date version of this document can be found at \url{https://github.com/HaysonC/skulenotes}\\

\section{More on Integrals}
\subsection{Riemann Sum - Non-Uniform Petition}
\begin{example}
    Given the following definite integral:
    $$\int^2_0 \sqrt{x} dx$$, we cannot evaluate its Riemann sum with uniforms partition, since the series of root cannot be easily evaluated. 
\end{example}

The definite integral of $\sqrt{x}$ from $0$ to $2$ using a Riemann sum with a non-uniform partition is given by:

$$\int_{0}^{2} \sqrt{x} \, dx = \lim_{n \to \infty} \sum_{i=1}^n \sqrt{x_i} \Delta x_i$$
, where:
\begin{itemize}
    \item $x_0 = 0, \, x_n = 2$,
    \item $x_i = i^2 \cdot \frac{2}{n^2}$ for $i = 0, 1, 2, \dots, n$,
    \item $\Delta x_i = x_i - x_{i-1} = \frac{2}{n^2} \cdot (2i - 1)$.
\end{itemize}

The Riemann sum becomes:
$$
S_n = \sum_{i=1}^n \sqrt{i^2 \cdot \frac{2}{n^2}} \cdot \frac{2}{n^2} \cdot (2i - 1).
$$

Simplifying further:
$$
S_n = \sum_{i=1}^n \sqrt{\frac{2i^2}{n^2}} \cdot \frac{2}{n^2} \cdot (2i - 1).
$$

Taking the limit as $n \to \infty$, the sum converges to the exact value of the integral using the series of squares. 
$$
\int_{0}^{2} \sqrt{x} \, dx = \frac{4\sqrt{2}}{3}.
$$
\paragraph{Condition} $n \to \infty$ Ensures $\Delta x_i \to 0$
\paragraph{} As $n \to \infty$, the partition points $x_i$ become increasingly dense. This ensures that the partition becomes infinitely fine.

\subsection{Integration By Parts}
Using the product rule:
$$
\frac{d}{dx} \big[ u(x) v(x) \big] = u'(x)v(x) + u(x)v'(x),
$$
integrating both sides with respect to $x$ gives:
$$
u(x)v(x) = \int u'(x) v(x) \, dx + \int u(x) v'(x) \, dx.
$$
Rearranging this:
$$\int u'(x) v(x) \, dx = u(x)v(x) - \int u(x) v'(x) \, dx.$$

\paragraph{Integration by parts formula}
\begin{equation}
\int u \, dv = uv - \int v \, du.
\end{equation}
\begin{example}
We want to solve the integral $$ \int x e^{2x} \, dx $$ using integration by parts.

Let:
$$ u = x, \quad dv = e^{2x} \, dx. $$

Then, we compute the derivatives and integrals:
$$ du = dx, \quad v = \frac{e^{2x}}{2}. $$

Now, apply the integration by parts formula:
$$ \int u \, dv = uv - \int v \, du. $$

Substituting in the values:
$$ \int x e^{2x} \, dx = x \cdot \frac{e^{2x}}{2} - \int \frac{e^{2x}}{2} \, dx. $$

Next, compute the remaining integral:
$$ \int \frac{e^{2x}}{2} \, dx = \frac{e^{2x}}{4}. $$

Thus, the result is:
$$ \int x e^{2x} \, dx = \frac{x e^{2x}}{2} - \frac{e^{2x}}{4} + C. $$

\end{example}

\begin{example}
We want to solve $$ \int x^2 \sin(2x) \, dx $$ using double integration by parts.

First, let:
$$ u = x^2, \quad dv = \sin(2x) \, dx. $$

Then:
$$ du = 2x \, dx, \quad v = -\frac{1}{2} \cos(2x). $$

Using the IBP formula:
$$ \int u \, dv = uv - \int v \, du, $$

we get:
$$ \int x^2 \sin(2x) \, dx = -\frac{x^2}{2} \cos(2x) + \int x \cos(2x) \, dx. $$

Now, apply IBP again to \( \int x \cos(2x) \, dx \), let:
$$ u = x, \quad dv = \cos(2x) \, dx. $$

Then:
$$ du = dx, \quad v = \frac{1}{2} \sin(2x). $$

Using the IBP formula again:
$$ \int x \cos(2x) \, dx = \frac{x}{2} \sin(2x) - \int \frac{1}{2} \sin(2x) \, dx, $$

and solving the remaining integral:
$$ \int \frac{1}{2} \sin(2x) \, dx = -\frac{1}{4} \cos(2x). $$

Thus, the final result is:
$$ \int x^2 \sin(2x) \, dx = -\frac{x^2}{2} \cos(2x) + \frac{x}{2} \sin(2x) + \frac{1}{4} \cos(2x) + C. $$

\end{example}
\subsection{Trigonometric Integrals}
\paragraph{Case I} This is the case I trigonometric integrals, the strategy is using the pythagorean identity.
\paragraph{} We want to solve the class of integrals:
\begin{equation}
 \int \sin^n(x) \cos^m(x) \, dx
\end{equation}
, where $m$ or $n$ is odd.
\begin{example}
We want to solve the integral:
$$ I = \int \sin^3(x) \cos^2(x) dx $$
We use the identity: $ \sin^2(x) = 1 - \cos^2(x)$. Thus, the integral becomes:
\begin{align*}
    I &= \int \sin(x)(1 - \cos^2(x)) \cos^2(x) \, dx \\
    &= \int (\cos^2(x) \sin(x) - \cos^4(x) \sin(x)) \, dx
\end{align*} 
which is now easily solvable with substution $u=\cos(x)$.
\end{example}
\paragraph{Case II} The follwoing is generally solvable via case I and case III below. In general, we solve the integral by reducing the power of the trigonometric functions to arrive at a solvable integral Case I or III.
\paragraph{} We want to solve the class of integrals:
\begin{equation} \int \sin^n(x) \cos^m(x) \, dx \end{equation}
, where $m$ and $n$ is even.
\begin{example}
We want to solve the integral:
$$ I = \int \sin^2(x) \cos^4(x) \, dx $$
We can apply the double angle formulas:
\begin{align}
    \sin(x)\cos(x) &= \frac{1}{2} \sin(2x) \\
    \sin^2(x) &= \frac{1 - \cos(2x)}{2} \\
    \cos^2(x) &= \frac{1 + \cos(2x)}{2}
\end{align}
Thus, the integral becomes:
$$ I = \frac{1}{8} \int \sin^2(2x) \, dx + \frac{1}{8} \int \cos(2x)\sin^2(2x) \, dx $$
\end{example}
\paragraph{Case III} This is the case III trigonometric integrals, the strategy is using the reduction formula via IBP.
\paragraph{Reduction Formula} We can solve integrals by reducing the power of the trigonometric functions. These are done using IBP and trigonometric identities.
\paragraph{} We want to solve the classes of integrals:
\begin{equation} 
    \int \sin^n(x) \, dx, \quad \int \cos^n(x) \, dx 
\end{equation}
, where $n$ is a positive integer. For demonstartion, We can obtain the reduction formula of $\sin^n(x)$ via IBP.
\begin{align}
    I_n &= \int \sin^n(x) \, dx \nonumber \\
    &= \int \sin^{n-1}(x) \sin(x) \, dx \nonumber \\
    &= \frac{-\cos(x) \sin^{n-1}(x)}{n} + \frac{(n-1)}{n} \int \cos^2(x) \sin^{n-2}(x) \, dx \nonumber \\
    &= \frac{-\cos(x) \sin^{n-1}(x)}{n} + \frac{(n-1)}{n} I_{n-2}
\end{align}
\paragraph{Reduction for $\cos^n(x)$} We can obtain the reduction formula of $\cos^n(x)$ via IBP.
\begin{align}
    I_n &= \int \cos^n(x) \, dx \nonumber \\
    &= \int \cos^{n-1}(x) \cos(x) \, dx \nonumber \\
    &= \frac{\sin(x) \cos^{n-1}(x)}{n} + \frac{(n-1)}{n} \int \sin^2(x) \cos^{n-2}(x) \, dx \nonumber \\
    &= \frac{\sin(x) \cos^{n-1}(x)}{n} + \frac{(n-1)}{n} I_{n-2}
\end{align} 
\paragraph{Case IV} The following is generally solvable via simple trigonometric integrals. In general, we solve the integral by applying the angle sum formulas.
\paragraph{} We want to solve the classes of integrals:
\begin{equation}
    \int \sin(mx) \cos(nx) \, dx, \quad \int \sin(mx) \sin(nx) \, dx, \quad \int \cos(mx) \cos(nx) \, dx 
\end{equation}
We could apply the angle sum formulas:
\begin{align}
    \sin(mx)\cos(nx) &= \frac{1}{2} \left[ \sin((m+n)x) + \sin((m-n)x) \right] \\
    \sin(mx)\sin(nx) &= \frac{1}{2} \left[ \cos((m-n)x) - \cos((m+n)x) \right] \\
    \cos(mx)\cos(nx) &= \frac{1}{2} \left[ \cos((m-n)x) + \cos((m+n)x) \right]
\end{align}
\paragraph{Case V} The following is generally solvable via the following trigonometric identities listed below, which convert it into a reduction formula.
\begin{align}
    \tan^2(x) &= \sec^2(x) - 1 \label{eq:tan2} \\
    \cot^2(x) &= \csc^2(x) - 1 \label{eq:cot2}  \\
    \frac{d}{dx} \tan(x) &= \sec^2(x)
\end{align}
\paragraph{} We want to solve the classes of integral:
\begin{equation}
    \int \tan^n(x) \, dx, \quad \int \cot^n(x) \, dx 
\end{equation}
We know that $tan^2(x) = \sec^2(x) - 1$. Thus, we can solve the integral by reducing the power of the tangent function.
\paragraph{Reduction for $\tan^n(x)$} We can obtain the reduction formula of $\tan^n(x)$ via the trigonometric identities.
\begin{align}
    I_n &= \int \tan^n(x) \, dx \nonumber \\
    &= \int \tan^{n-1}(x) \tan(x) \, dx \nonumber \\
    &= \frac{\tan^{n-1}(x)}{n-1} - \int \tan^{n-2}(x) \, dx \nonumber \\
    &= \frac{\tan^{n-1}(x)}{n-1} - I_{n-2}
\end{align}
\paragraph{Reduction for $\cot^n(x)$} We can obtain the reduction formula of $\cot^n(x)$ via the trigonometric identities.
\begin{align}
    I_n &= \int \cot^n(x) \, dx \nonumber \\
    &= \int \cot^{n-1}(x) \cot(x) \, dx \nonumber \\
    &= \frac{\cot^{n-1}(x)}{n-1} - \int \cot^{n-2}(x) \, dx \nonumber \\
    &= \frac{\cot^{n-1}(x)}{n-1} - I_{n-2}
\end{align}
\paragraph{Case VI} The following is generally solvable via the trigonometric identities (\ref{eq:tan2}) and (\ref{eq:cot2}).
\paragraph{} We want to solve the classes of integral:
\begin{equation} 
    \int \tan^m(x) \sec^n(x) \, dx, \quad \int \cot^m(x) \csc^n(x) \, dx 
\end{equation}
We can solve the integral by reducing the power of the converted trigonometric functions using (\ref{eq:tan2}) and (\ref{eq:cot2}).
\paragraph{Case VII} The following is generally solvable via the trigonometric identities (\ref{eq:tan2}) and (\ref{eq:cot2}).
\paragraph{} We want to solve the classes of integral:
\begin{equation} \int \tan^m(x) \sec^n(x) \, dx, \quad \int \cot^m(x) \csc^n(x) \, dx \end{equation}
We can solve the integral by converting between the tangent and secant functions using the trigonometric identities.
\paragraph{Case VIII} This is the case VIII integrals, the strategy is using the trigonometric substitution.
\paragraph{Trigonometric Substitution} We can solve integrals by substituting the trigonometric functions with other trigonometric functions.
\begin{example}
We want to solve the integral:
$$ \int \frac{dx}{\sqrt{1 - x^2}} $$
\end{example}
We can substitute $x = \sin(\theta)$, then $dx = \cos(\theta) \, d\theta$. The integral becomes:
$$ \int \frac{\cos(\theta) \, d\theta}{\sqrt{1 - \sin^2(\theta)}} = \int \frac{\cos(\theta) \, d\theta}{\cos(\theta)} = \int d\theta = \theta + C. $$
\paragraph{} In general, we can use the following substitutions:
\begin{itemize}
    \item $x = a\sin(\theta)$ for $\sqrt{a^2 - x^2}$,
    \item $x = a\tan(\theta)$ for $\sqrt{a^2 + x^2}$,
    \item $x = a\sec(\theta)$ for $\sqrt{x^2 - a^2}$.
\end{itemize}
\paragraph{Weierstrass Substitution} We can solve integrals by using the Weierstrass substitution:
\begin{equation}
    \tan\left(\frac{x}{2}\right) = t \Leftrightarrow \cos(x) = \frac{1-t^2}{1+t^2} \quad \sin(x) = \frac{2t}{1+t^2} \quad dx = \frac{2dt}{1+t^2}
\end{equation}
This substitution is useful for solving integrals with trigonometric functions, by converting them into rational functions.
\paragraph{Summary} We can solve trigonometric integrals by using the following strategies:
\begin{table}[h!]
    \centering
    \begin{tabular}{|c|l|}
    \hline
    \textbf{Case} & \textbf{Strategy and General Form} \\ \hline
    I   & Use \(\sin^2(x) + \cos^2(x) = 1\); simplify using substitution. \\
        & General Form: \(\int \sin^n(x) \cos^m(x) dx\), where \(m\) or \(n\) is odd. \\ \hline
    II  & Convert to Case I and Case III using double angle formulas. \\
        & General Form: \(\int \sin^n(x) \cos^m(x) dx\), where both \(m\) and \(n\) are even. \\ \hline
    III & Apply reduction formulas derived via integration by parts. \\
        & General Form: \(\int \sin^n(x) dx\) or \(\int \cos^n(x) dx\). \\ \hline
    IV  & Use angle sum formulas to simplify. \\
        & General Form: \(\int \sin(mx) \cos(nx) dx\). \\ \hline
    V   & Reduce tangent/cotangent powers using \(\tan^2(x) = \sec^2(x) - 1\) and substitution. \\
        & General Form: \(\int \tan^n(x) dx\) or \(\int \cot^n(x) dx\). \\ \hline
    VI  & Convert to Case V via Pythagorean identities. \\
        & General Form: \(\int \tan^m(x) \sec^n(x) dx\) or \(\int \cot^m(x) \csc^n(x) dx\). \\ \hline
    VII & Convert between tangent and secant functions for simplification. \\
        & General Form: \(\int \tan^m(x) \sec^n(x) dx\). \\ \hline
    VIII & Use trigonometric substitution: \(x = a \sin(\theta), a \tan(\theta), a \sec(\theta)\). \\
         & General Form: \(\int \frac{dx}{\sqrt{a^2 - x^2}}, \int \frac{dx}{\sqrt{a^2 + x^2}}, \int \frac{dx}{\sqrt{x^2 - a^2}}\). \\ \hline
    \end{tabular}
    \caption{Strategies and General Forms for Case I-VIII Integrals}
\end{table}
\subsection{Partial Fractions}
\paragraph{Partial Fraction Decomposition} We can decompose a rational function into partial fractions to simplify integration.
\paragraph{} We want to solve the integral:
\begin{equation}
    \int \frac{P(x)}{Q(x)} \, dx
\end{equation}
, where $P(x)$ and $Q(x)$ are polynomials.
\begin{example}
We want to solve the integral:
$$ \int \frac{2x^2 + 3x + 1}{x^3 + 2x^2 + x} \, dx $$
\end{example}
We can decompose the rational function into partial fractions:
$$ \frac{2x^2 + 3x + 1}{x^3 + 2x^2 + x} = \frac{A}{x} + \frac{B}{x+1} + \frac{C}{(x+1)^2} $$
\paragraph{} We can solve for the constants $A$, $B$, and $C$ by equating the coefficients of the partial fractions to the original function. We have:
\begin{align*}
    2x^2 + 3x + 1 &= A(x+1)^2 + Bx(x+1) + Cx \\
    \intertext{Thus, we can solve it by setting $x = 0$, $x = -1$}
\end{align*}
\paragraph{Summary} We can solve integrals using partial fraction decomposition by following these steps:
\begin{table}[h!]
    \centering
    \begin{tabular}{|c|l|}
    \hline
    \textbf{Step} & \textbf{Description} \\ \hline
    1 & Factorize the denominator of the rational function. \\
    2 & Write the partial fraction decomposition. \\
    3 & Solve for the constants by equating the coefficients. \\
    4 & Integrate the partial fractions. \\ \hline
    \end{tabular}
    \caption{Steps for Partial Fraction Decomposition}
\end{table}
\subsection{Improper Integrals}
\paragraph{Improper Integral} An improper integral is an integral with an infinite limit or a discontinuity in the interval of integration.
\paragraph{} We want to solve the integral:
\begin{equation}
    \int_a^b f(x) \, dx
\end{equation}
, where $a$ or $b$ is infinite or $f(x)$ is discontinuous.
\begin{example}
We want to solve the integral:
$$ \int_0^\infty e^{-x} \, dx $$

Evaluating the integral:
\begin{align*}
    \int_0^\infty e^{-x} \, dx &= \lim_{b \to \infty} \int_0^b e^{-x} \, dx \\
    &= \lim_{b \to \infty} -e^{-x} \Big|_0^b \\
    &= \lim_{b \to \infty} -e^{-b} + 1 \\
    &= 1
\end{align*}
\end{example}
\subsection{Convergence Test}
\paragraph{Convergence} An improper integral converges if the limit of the integral exists. 
\paragraph{Comparison Test} We can compare the integral to another integral to determine convergence or divergence.
\paragraph{Spliting the Integral} We can split the integral into two parts to determine convergence. We split it into two limits on the left and right, for example:
$$ \int_a^b f(x)\, dx = \lim_{t \to c^-} \int_a^t f(x) \, dx + \lim_{t \to c^+} \int_t^b f(x) \, dx $$
\section{Hyperbolic Trigonometric Functions}
\begin{definition}[Hyperbolic Sine]
    The hyperbolic sine function is defined as:
    \begin{equation} \sinh(x) = \frac{e^x - e^{-x}}{2}. \end{equation}
\end{definition}
\begin{definition}[Hyperbolic Cosine]
    The hyperbolic cosine function is defined as:
    \begin{equation}  \cosh(x) = \frac{e^x + e^{-x}}{2}. \end{equation}
\end{definition}
\paragraph{Properties} These combinations of exponential functions have properties similar to the trigonometric functions.
\paragraph{Derivatives} The derivatives of the hyperbolic functions are:
\begin{align}
    \frac{d}{dx} \sinh(x) &= \cosh(x), \\
    \frac{d}{dx} \cosh(x) &= \sinh(x).
\end{align}
\paragraph{Identities} The hyperbolic functions satisfy the following identities:
\begin{align}
    \cosh^2(x) - \sinh^2(x) &= 1, \\
    \cosh(2x) &= \cosh^2(x) + \sinh^2(x), \\
    \sinh(2x) &= 2\sinh(x)\cosh(x).
\end{align}
\paragraph{Hyperbola} The hyperbolic functions are related to the hyperbola $x^2 - y^2 = 1$. Simiar to the circle, the hyperbola can be parametrized by the hyperbolic functions (e.g. $x = \cosh(t)$, $y = \sinh(t)$).
\paragraph{Area} The area of a sector of the hyperbola is given by:
\begin{equation}
    A = t/2,
\end{equation} 
where $t$ is the angle of the sector along the parametrization $\{(x,y)\mid x = \cosh(t), y = \sinh(t)\}$.
\begin{figure}[h!]
    \begin{center}
        \begin{tikzpicture}[scale=0.8]
            % Draw the axes
            \draw[->] (-5, 0) -- (5, 0) node[right] {$x$};
            \draw[->] (0, -5) -- (0, 5) node[above] {$y$};

            % Draw ticks on the axes
            \foreach \x in {-4, -2, 2, 4}
                \draw (\x, 0.1) -- (\x, -0.1) node[below] {\x};
            \foreach \y in {-4, -2, 2, 4}
                \draw (0.1, \y) -- (-0.1, \y) node[left] {\y};

            % Draw the hyperbola branches
            \draw[thick, red, domain=-2:2, samples=100] plot ({cosh(\x)}, {sinh(\x)});
            \draw[thick, red, domain=-2:2, samples=100] plot ({-cosh(\x)}, {sinh(\x)});

            % Draw the asymptotes
            \draw[red, dashed] (-5, -5) -- (5, 5);
            \draw[red, dashed] (-5, 5) -- (5, -5);

            % Add a point on the hyperbola
            \coordinate (P) at ({cosh(1)}, {sinh(1)});
            \draw[fill] (P) circle (0.05) node[above right] {$(\cosh(t), \sinh(t))$};

            % Connect the labeled point to the origin
            \draw[thick, blue] (0, 0) -- (P);

            % Fill the sector area between hyperbola, line, and x-axis
            \fill[blue, opacity=0.3, domain=0:1, variable=\t]
                (0, 0) -- plot ({cosh(\t)}, {sinh(\t)}) -- cycle;
        \end{tikzpicture}
    \end{center}
    \caption{Sector of the hyperbola}
    \centering
\end{figure}
\paragraph{Catenary} The hyperbolic cosine function describes the shape of a hanging chain or cable. The catenary is the curve formed by a chain hanging from two points. It is given by the equation:
\begin{equation}
    a \cosh\left(\frac{x}{a}\right),
\end{equation}
\begin{definition}[Hyperbolic Tangent]
    The hyperbolic tangent function is defined as:
    \begin{equation}
        \tanh(x) = \frac{\sinh(x)}{\cosh(x)} = \frac{e^x - e^{-x}}{e^x + e^{-x}}. 
    \end{equation}
\end{definition}
\paragraph{Derivative} The derivative of the hyperbolic tangent function resembles the derivative of the regular tangent function:
\begin{equation} \frac{d}{dx} \tanh(x) = \text{sech}^2(x). \end{equation}
\paragraph{Identities} The hyperbolic tangent function satisfies the following identities:
\begin{align}
    \tanh(x) &= \frac{\sinh(x)}{\cosh(x)}, \\
    \text{sech}^2(x) &= 1 - \tanh^2(x).
\end{align}
\paragraph{Secant, Cosecant, Cotangent} The hyperbolic secant, cosecant, and cotangent functions are defined similarly to the regular secant, cosecant, and cotangent functions. They are the \textbf{reciprocal} of the hyperbolic cosine, sine, and tangent functions, respectively.
\paragraph{Inverse Hyperbolic Functions} The inverse hyperbolic functions are defined as the inverse of the hyperbolic functions. They are denoted by $\sinh^{-1}(x)$, $\cosh^{-1}(x)$, $\tanh^{-1}(x)$, etc.
\begin{align}
    \text{arsinh } x &= \ln\left(x + \sqrt{x^2 + 1}\right) \\[10pt]
    \text{arcosh } x &= \ln\left(x + \sqrt{x^2 - 1}\right) \\[10pt]
    \text{artanh } x &= \frac{1}{2} \ln\left(\frac{1 + x}{1 - x}\right) \\[10pt]
    \text{arcsch } x &= \ln\left(\frac{1}{x} + \sqrt{\frac{1}{x^2} + 1}\right) \\[10pt]
    \text{arsech } x &= \ln\left(\frac{1}{x} + \sqrt{\frac{1}{x^2} - 1}\right) \\[10pt]
    \text{arcoth } x &= \frac{1}{2} \ln\left(\frac{x + 1}{x - 1}\right)
\end{align}
\section{Further Applications of Integrals}
\subsection{Arc Length}
\paragraph{Arc Length}
    Given the curve $C: y = f(x)$, the arc length of the curve between two points $a$ and $b$, where $y'(x)$ is continuous, we would like the compute the arc length of the curve:
    $$ L = \int_C ds$$
\paragraph{Formula} We can derive the formula for the arc length of a curve $y = f(x)$ between two points $a$ and $b$ as:
\begin{align}
    \intertext{At $x_i$ and $x_{i+1}$, the length of the segment is:}
    \Delta s &\approx \sqrt{(x_{i+1} - x_i)^2 + (y_{i+1} - y_i)^2} \nonumber \\
            &= \sqrt{\Delta_i x^2 + \Delta_i y^2} \nonumber \\
    \intertext{We also have, by MVT:}
    \frac{\Delta y}{\Delta x} &= y'(c) \nonumber \\
    \Delta y &= y'(c) \Delta x \nonumber \\
    \Delta s &\approx \sqrt{\Delta x^2 + (y'(c) \Delta x)^2} \nonumber \\
            &= \sqrt{1 + y'(c)^2} \Delta x \nonumber \\
    \intertext{Taking a Riemann sum:}
    L &= \lim_{n \to \infty} \sum_{i=1}^n \sqrt{1 + y'(c_i)^2} \Delta x \nonumber \\
    &= \int_a^b \sqrt{1 + y'(x)^2} \, dx 
\end{align}
\begin{example}
    Given $f(x) = x^\frac{3}{2}$ between $x = 0$ and $x = 44$, we would like to compute the arc length of the curve:
    \begin{align*}
        L &= \int_0^{44} \sqrt{1 + \left(\frac{3}{2}x^{\frac{1}{2}}\right)^2} \, dx \\
        &= \int_0^{44} \sqrt{1 + \frac{9}{4}x} \, dx \\
        &= \frac{8}{27} \left(1 + \frac{9}{4}x\right)^{\frac{3}{2}} \Big|_0^{44} = 296
    \end{align*}
\end{example}
\paragraph{Inverse} Also, we can take $x = f^{-1}(y)$ to compute the arc length of a curve.
\begin{equation}
    L = \int_{f(a)}^{f(b)} \sqrt{1 + ((f{^{-1}})'(y))^2} \, dy
\end{equation}
\subsection{Surface Area of Revolution}
\paragraph{Surface Area of Revolution} Given the curve $C: y = f(x)$, we would like to compute the surface area of the curve between two points $a$ and $b$ when rotated about the $x$-axis:
    $$ A_x = \int_C 2\pi y \, ds$$
\paragraph{Formula} Deriving the formula for the surface area of a curve $y = f(x)$ between two points $a$ and $b$ rotated about the $x$-axis as follows:
\begin{align}
    \intertext{At $x_i$ and $x_{i+1}$, the length of the segment is:}
    \Delta s &\approx \sqrt{\Delta_i x^2 + \Delta_i y^2} \nonumber \\
    \intertext{The surface area of the segment is, by continuity:}
    A_i &= \pi (y_i + y_{i+1}) \Delta s = 2\pi y^\star \Delta s \quad 
    \intertext{As the difference between $y^\star$ and $y_i$ diminishes as $\Delta x \rightarrow 0$, taking a Riemann sum:}
    A_x &= \lim_{n \to \infty} \sum_{i=1}^n 2\pi y_i \sqrt{\Delta_i x^2 + \Delta_i y^2} \nonumber \\
    &= \int_a^b 2\pi y \sqrt{1 + y'(x)^2} \, dx
    \intertext{The surface area of a curve $y = f(x)$ (i.e. $x = f^{-1}$)between two points $f(a)$ and $f(b)$ rotated about the $y$-axis is:}
    A_y &= \int_{f(a)}^{f(b)} 2\pi x \sqrt{1 + x'(y)^2} \, dy
\end{align}
\begin{example}
    Given $y = \sqrt{x}$ between $x = 0$ and $x = 1$, we would like to compute the surface area of the curve:
    \begin{align*}
        A_x &= \int_0^1 2\pi \sqrt{x} \sqrt{1 + \frac{1}{4x}} \, dx \\
        &= \pi \int^1_0 \sqrt{4x + 1} \, dx \\
        \intertext{Let $u = 4x + 1$, then $du = 4dx$:}
        &= \frac{\pi}{4} \int^5_1 \sqrt{u} \, du \\
        &= \frac{\pi}{6}\left(5^\frac{3}{2} -1\right)
    \end{align*}
\end{example}
\subsection{Applications to Physics and Engineering}
\subsubsection{Hydrostatic Pressure and Force}
\paragraph{Hydrostatic Pressure} The hydrostatic pressure and Force is given by the Archimedes' principle:
\begin{subequations}
    \begin{align}
        P &= \rho g \cdot d, \\
        F &= \rho g \cdot d \cdot A
    \end{align}
\end{subequations}
\paragraph{Riemann Sum} Given a tank rectangular with width defined by $w = f(x)$, we can take a horizontal slice, $x^\star$, of the tank to compute the force exerted on the tank by the water:
\begin{equation}
    F =\sum_{i=1}^n \rho g x_i f(x_i) \Delta x
\end{equation}
\paragraph{Formula} Given a tank rectangular with width defined by $w = f(x)$, where $x$ is depth, we would like to compute the force exerted on the tank by the water:
\begin{equation}
    F = \int_0^b \rho g x f(x) \cdot dx
\end{equation}
, which is derived by considering $dx$ as the depth a horizontal slice of the tank.
\subsubsection{Center of Mass and Moments of Inertia}
\paragraph{Properties of the Center of Mass} The center of mass of a region $R$ has the following properties:
\begin{enumerate}
    \item \textbf{Symetry}: For all axis of symmetry, the center of mass lies on the axis.
    \item \textbf{Additivity}: The center of mass of a region is the weighted average of the centers of mass of its parts. The weights are the areas of the parts.
    \begin{equation}
        \bar{x} = \sum_i \frac{A_i\bar{x_i}}{A}
    \end{equation}
\end{enumerate}
\paragraph{Formula} The centroid of a uniformly-dense region $R$ bounded by $y=f(x),\,x\in[a,b],\,y\in[f(a),f(b)]$ is derived as follows:
\begin{align}
    \intertext{We first have $x_i^\star$ and $A_i$ as follows:}
    x_i^\star = \frac{1}{2}(x_i + x_{i+1}) &\quad A_i = f(x_i) \Delta x \nonumber \\
    \intertext{For $\bar{x}$, we have:}
    A\bar{x} &\approx \sum_i x_i^\star A_i \nonumber \\
    \intertext{Taking a Riemann sum:}
    A\bar{x} &= \lim_{n \to \infty} \sum_i x_i^\star A_i  \nonumber \\
    \bar{x} &= \frac{1}{A}\int_a^b x f(x) \, dx
    \intertext{Similarly, for $\bar{y}$:}
    A\bar{y} &= \lim_{n \to \infty} \sum_i f(x_i^\star) A_i \nonumber \\
    \bar{y} &= \frac{1}{A}\int_a^b \frac{1}{2}(f(x))^2 \, dx
\end{align}
\begin{theorem}[Pappus's Theorem]
    The volume of a solid of revolution is given by the product of the area of the region and the distance, $R$, traveled by the centroid of the region from the axis of rotation.
    \begin{equation}
        V = 2\pi R A
    \end{equation}
\end{theorem}
\begin{example}[Eliptical Torus]
    Given an eliptical torus (tall donut) with major axis $a$ and minor axis $b$, we would like to compute the volume of the torus:
    \begin{align*}
        V &= 2\pi R A \\
        &= 2\pi R \cdot \pi ab = 2\pi^2 abR
    \end{align*}
    This could be, alternatively, shown via the washer method and the shell method.
\end{example}
\subsection{Applications to Economics and Biology}
``Biologists are so bad at math'' - Prof. Davis as he proceeds to erase the board and skip to the next section.
\section{Parametric Equations and Polar Coordinates}
\subsection{Curves Defined by Parametric Equations}
\paragraph{Parametric Equations} For $t \in \mathbb{R}$, we have:
\begin{equation}
    x = x(t), \quad y = y(t)
\end{equation}
\begin{example}[Newton's Laws of Motion]
    Given $\ddot{x} = 0$, $\ddot{y} = -g$, we have:
    \begin{align*}
        x(t) &= A_1 t + A_2, \\
        y(t) &= -\frac{1}{2}gt^2 + B_1 t + B_2
    \end{align*}
    At initial conditions $t = 0$, we have $x(0) = 0$, $y(0) = 0$, $\dot x(0) = v_0 \cos(\theta)$, $\dot y(0) = v_0 \sin(\theta) - \frac{1}{2}gt^2$.
\end{example}
\begin{example}[Ellipse]
    Given $x = a\cos(t)$, $y = b\sin(t)$, we have:
    \begin{align*}
        x^2 &= a^2\cos^2(t), \\
        y^2 &= b^2\sin^2(t)
    \end{align*}
    , which satisfies $x^2/a^2 + y^2/b^2 = 1$.
\end{example}
\paragraph{Intersections of Parametric Curves} Given two parametric curves $C_1,\,C_2$, defined by:
\begin{align*}
    C_1:\, x_1 &= x_1(t), \quad y_1 = y_1(t) \\
    C_2:\, x_2 &= x_2(t), \quad y_2 = y_2(t)
\end{align*}
, we can find the intersection points by solving the system of equations:
\begin{align*}
    x_1(t) &= x_2(t) \\
    y_1(t) &= y_2(t)
\end{align*}
\begin{example}
    Given $x_1 = 2t+t$, $y_1 = 5-4t$, $x_2 = 3 - 5\cos(\pi t)$, $y_2 = 1 + 5\sin(\pi t)$, we have:
    \begin{align*}
        2t + 1 &= 3 - 5\cos(\pi t) \\
        5 - 4t &= 1 + 5\sin(\pi t)
    \end{align*}
\end{example}
\subsection{Calculus with Parametric Curves}
\paragraph{Tangents} The tangent to a parametric curve $C$ at a point $t$ is given by:
\begin{equation}
    \frac{dy}{dx} = \frac{dy/dt}{dx/dt} = \frac{y'(t)}{x'(t)}
\end{equation}
\paragraph{Equation of the Tangent} The equation of the tangent to a parametric curve $C$ at a point $t_0$ is given by:
\begin{equation}
    y'(t_0) (x-x(t_0)) - x'(t_0) (y-y(t_0)) = 0
\end{equation}
Hence, if $x'(t_0) = 0$, we have a vertical tangent; if $y'(t_0) = 0$, we have a horizontal tangent. However, if $x'(t_0) = 0$ and $y'(t_0) = 0$, we have no information.
\begin{example}
    Let $x = \sin 2t$, $y = \sin t$, we have:
    \begin{align*}
        x'(t) &= 2\cos 2t, \\
        y'(t) &= \cos t
    \end{align*}
    \textbf{Vertical Tangent}: $2\cos 2t = 0 \Rightarrow t = \frac{\pi}{4} + \frac{n\pi}{2}$.
    \textbf{Horizontal Tangent}: $\cos t = 0 \Rightarrow t = \frac{\pi}{2} + n\pi$.
    \textbf{At $t = 0$} we have $x' = y' = 0$. Thus, we have no information. On the graph, we see that the curve intersects itself at $t = 0$.
\end{example}
\paragraph{Area under Parametric Curve} The area under a parametric curve $C$ between $t_1$ and $t_2$ is given by:
\begin{equation}
    A = \int^{x(t_2)}_{x(t_1)} y(x) \, dx = \int_{t_1}^{t_2} y(t) x'(t) \, dt 
\end{equation}
\begin{definition}[Orientation]
    The orientation of a parametric curve $C$ is given by the direction of the curve as $t$ increases. If the enclosed area is to the left of the trace of $t$ (counter-clockwise) the orientation is positive; otherwise, it is negative.
\end{definition}
\paragraph{} The sign of the area of a parametric curve $C$ is given by the orientation of the curve.
\paragraph{Area of Closed Curves} The area of a closed parametric curve $C$ is given by:
\begin{equation}
    A = \oint_C dA = \int_{t_1}^{t_2} y(t) x'(t) \, dt = \int_{t_1}^{t_2} x(t) y'(t) \, dt
\end{equation}
, where
\begin{equation*}
    x(t_1) = x(t_2), \quad y(t_1) = y(t_2) \quad \text{and} \quad \text{$t_2$ is the smallest $t>t_1$ to satisfy the condition}.
\end{equation*}
\paragraph{Arc Length} The arc length of a parametric curve $C$ is given by:
\begin{equation}
    L = \int_a^b \sqrt{\left(\frac{dx}{dt}\right)^2 + \left(\frac{dy}{dt}\right)^2} \, dt
\end{equation}
\begin{example}
    Given $x = t\cos t$, $y = t\sin t$, we have:
    \begin{align*}
        \frac{dx}{dt} &= \cos t - t\sin t, \\
        \frac{dy}{dt} &= \sin t + t\cos t
    \end{align*}
    Thus, the arc length is given by:
    \begin{align*}
        L &= \int_0^{2\pi} \sqrt{(\cos t - t\sin t)^2 + (\sin t + t\cos t)^2} \, dt \\
        &= \int_0^{2\pi} \sqrt{1 + t^2} \, dt
        \intertext{Using the substitution $t = \tan \theta$, we have:}
        &= \pi \sqrt{1 + 4\pi^2} + \frac{1}{2}\ln(2\pi + \sqrt{1 + 4\pi^2})
    \end{align*}
\end{example}
\paragraph{Surface Area of Revolution} The surface area of a parametric curve $C$ rotated about the $x$-axis is given by:
\begin{align}
    A &= \int_a^b 2\pi y(t) ds \nonumber \\
    \intertext{From the arc length formula, we have $ds = \sqrt{(\frac{dx}{dt})^2 + (\frac{dy}{dt})^2} \, dt$, thus:}
    &= \int_a^b 2\pi y(t) \sqrt{\left(\frac{dx}{dt}\right)^2 + \left(\frac{dy}{dt}\right)^2} \, dt
\end{align}
Similarly, the surface area of a parametric curve $C$ rotated about the $y$-axis is given by:
\begin{equation}
    A = \int_a^b 2\pi x(t) \sqrt{\left(\frac{dx}{dt}\right)^2 + \left(\frac{dy}{dt}\right)^2} \, dt
\end{equation}
\begin{example}[Surface Area of of a Ellipse]
    Given $x = a\sin t$, $y = b\cos t$, we have:
    \begin{align*}
        \frac{dx}{dt} &= a\cos t, \\
        \frac{dy}{dt} &= -b\sin t
    \end{align*}
    Thus, the surface area of the ellipse is given by:
    \begin{align*}
        A &= \int_0^{2\pi} 2\pi b\cos t \sqrt{a^2\cos^2 t + b^2\sin^2 t} \, dt \\
        &= 2\pi b \int_0^{2\pi} \sqrt{a^2 (1 - \sin^2 t) + b^2 \sin^2 t} \, dt \\
        &= 2\pi b \int_0^{2\pi} \sqrt{a^2 + b^2 - (a^2 - b^2)\sin^2 t} \, dt
    \end{align*}
    This has no analytical solution, but can be solved numerically. Typically, we take $\epsilon = \sqrt{\frac{a^2-b^2}{a^2}}$.
\end{example}
\section{Polar Coordinates}
\paragraph{Polar Coordinates} Given a point $P$ in the plane, we can define the polar coordinates of $P$ as $(r, \theta)$, where $r$ is the distance from the origin to $P$ and $\theta$ is the angle between the positive $x$-axis and the line segment from the origin to $P$.
\paragraph{Transformation} We can convert between polar and Cartesian coordinates as follows:
\begin{equation}
    x = r\cos \theta, \quad y = r\sin \theta
\end{equation}
\paragraph{Reverse Transformation} We can convert between Cartesian and polar coordinates as follows:
\begin{equation}
    r = \sqrt{x^2 + y^2}, \quad \theta = \arctan\left(\frac{y}{x}\right)
\end{equation}
\paragraph{Note} The angle $\theta$ is not unique, as $\theta + 2\pi n$ for $n \in \mathbb{Z}$ represents the same point, beware of this when converting between polar and Cartesian coordinates.
\begin{example}[Lines]
    \begin{align*}
        y = mx + b &\implies \theta = \alpha = \arctan m \\
        x = a &\implies r = a\sec \theta \\
        y = a &\implies r = a\csc \theta
    \end{align*}
\end{example}
\begin{example}[Cirlces]
    \begin{align*}
        x^2 + y^2 = a^2 &\implies r = a \\
    \end{align*}
    Let $r = 6\sin \theta$, we have:
    \begin{align*}
        r^2 &= 6r\sin \theta \\
        x^2 + y^2 &= 6y \\
        x^2 + (y-3)^2 &= 9
    \end{align*}
    We deduce that the curve is a circle with radius $3$ and center $(0,3)$.
\end{example}
\paragraph{Cylinrical Coordinates} Given a point $P$ in space, we can define the cylindrical coordinates of $P$ as $(r, \theta, z)$, where $r$ is the distance from the $z$-axis to $P$, $\theta$ is the angle between the positive $x$-axis and the projection of the line segment from the origin to $P$ onto the $xy$-plane, and $z$ is the distance from the $xy$-plane to $P$.
\paragraph{Spherical Coordinates} Given a point $P$ in space, we can define the spherical coordinates of $P$ as $(\rho, \theta, \phi)$, where $\rho$ is the distance from the origin to $P$, $\theta$ is the angle between the positive $x$-axis and the projection of the line segment from the origin to $P$ onto the $xy$-plane, and $\phi$ is the angle between the positive $z$-axis and the line segment from the origin to $P$.
\subsection{Graphing in Polar Coordinates}
\begin{example}
    Let $r = \frac{1}{2} + \cos \theta$, we first figure out the origin of the curve by setting $r = 0$:
    \begin{align*}
        0 &= \frac{1}{2} + \cos \theta \\
        \theta &= \frac{2\pi}{3},\, \frac{4\pi}{3}
    \end{align*}
    We then deduce the maximum and minimum of the curve to the origin by setting $\frac{dr}{d\theta} = 0$:
    \begin{align*}
        \frac{dr}{d\theta} &= -\sin \theta = 0 \\
        \theta &= 0,\, \pi,\, 2\pi
    \end{align*}
    We also look for symmetry in the curve by checking $r(\theta) = r(-\theta)$:
    \begin{align*}
        \frac{1}{2} + \cos \theta &= \frac{1}{2} + \cos(-\theta) \\
    \end{align*}
    Thus, the curve is symmetric about the $x$-axis. We also look for symmetry about the $y$-axis by checking $r(\theta) = r(\pi + \theta)$:
    \begin{align*}
        \frac{1}{2} + \cos \theta \neq \frac{1}{2} + \cos(\pi + \theta) \\
    \end{align*}
    Thus, the curve is not symmetric about the $y$-axis. We also look for intervals in which the curve is increasing or decreasing by checking $r'(\theta) > 0$ or $r'(\theta) < 0$.
\end{example}
\paragraph{Common Polar Curves} Some common polar curves are:
\begin{enumerate}
    \item \textbf{Cardioid}: $r = a(1 + \cos \theta)$ (heart-shaped)
    \item \textbf{Circle}: $r = a\cos \theta$ 
    \item \textbf{Limasçons}: $r = a + b\sin \theta$ (heart with a hole)
    \item \textbf{Lemniscate}: $r^2 = a^2\sin 2\theta$ (infinity symbol, diagonal) \vspace{1em} $\quad r^2 = a^2\cos 2\theta$ (infinity symbol, horizontal)
    \item \textbf{Petal Curves}: $r = a\cos n\theta$ ($mn$-petal flower) \vspace{1em} $\quad r = a\sin n\theta$ ($mn$-petal flower); $m=2$ if $n$ is even, $m=1$ if $n$ is odd.
\end{enumerate}
\paragraph{} Below are some examples of polar curves:
\begin{figure}[h!]
    \centering
    \begin{minipage}{0.48\textwidth}
        \centering
        \begin{tikzpicture}[scale=1.5]
            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
            \draw[domain=0:360, smooth, samples=100, red] plot ({(1+cos(\x))*cos(\x)}, {(1+cos(\x))*sin(\x)});
            \node at (0,-3) {Cardioid: $r = 1 + \cos \theta$};
        \end{tikzpicture}
    \end{minipage}
    \hfill
    \begin{minipage}{0.48\textwidth}
        \centering
        \begin{tikzpicture}[scale=1.5]
            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
            \draw[domain=0:360, smooth, samples=100, blue] plot ({cos(\x)*cos(\x)}, {cos(\x)*sin(\x)});
            \node at (0,-3) {Circle: $r = \cos \theta$};
        \end{tikzpicture}
    \end{minipage}
\end{figure}

\begin{figure}[h!]
    \centering
    \begin{minipage}{0.48\textwidth}
        \centering
        \begin{tikzpicture}[scale=1.5]
            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
            \draw[domain=0:360, smooth, samples=100, green] plot ({(1+0.5*sin(\x))*cos(\x)}, {(1+0.5*sin(\x))*sin(\x)});
            \node at (0,-3) {Limaçon: $r = 1 + 0.5\sin \theta$};
        \end{tikzpicture}
    \end{minipage}
    \hfill
    \begin{minipage}{0.48\textwidth}
        \centering
        \begin{tikzpicture}[scale=1.5]
            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
            \draw[domain=0:360, smooth, samples=100, orange] plot ({sqrt(abs(sin(2*\x)))*cos(\x)}, {sqrt(abs(sin(2*\x)))*sin(\x)});
            \node at (0,-3) {Lemniscate: $r^2 = \sin 2\theta$};
        \end{tikzpicture}
    \end{minipage}
\end{figure}

\begin{figure}[h!]
    \centering
    \begin{minipage}{0.48\textwidth}
        \centering
        \begin{tikzpicture}[scale=1.5]
            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
            \draw[domain=0:360, smooth, samples=100, purple] plot ({cos(3*\x)*cos(\x)}, {cos(3*\x)*sin(\x)});
            \node at (0,-3) {Petal Curve: $r = \cos 3\theta$};
        \end{tikzpicture}
    \end{minipage}
    \hfill
    \begin{minipage}{0.48\textwidth}
        \centering
        \begin{tikzpicture}[scale=1.5]
            \draw[->] (-2.5,0) -- (2.5,0) node[right]{$x$};
            \draw[->] (0,-2.5) -- (0,2.5) node[above]{$y$};
            \draw[domain=0:360, smooth, samples=100, brown] plot ({sin(2*\x)*cos(\x)}, {sin(2*\x)*sin(\x)});
            \node at (0,-3) {Petal Curve: $r = \sin 2\theta$};
        \end{tikzpicture}
    \end{minipage}
\end{figure}
\newpage
\paragraph{} \label{sec:polar} A detailed list of polar curves and their properties could be found in Appendix \ref{app:polarCurves}. 
\subsection{The intersection of Polar Curves}
\begin{example}
    Let $r_1 = \sin \theta$, $r_2 = -\cos \theta$, we have:
    \begin{align*}
        \sin \theta &= -\cos \theta \\
        \tan \theta &= -1 \\
        \theta &= \frac{3\pi}{4},\, \frac{7\pi}{4}
    \end{align*}
    At these points, we can find the $x$ and $y$ coordinates of the intersection points. But also, we must check for the origin point $(x,y) = (0,0)$. Also $r_1 = r_2$ is not always reliable so we have to check it intuitively.
\end{example}
\subsection{Area and Length in Polar Coordinates}
\paragraph{Area in Polar Coordinates} The area of a region $R: r = g(\theta)$ in polar coordinates is given by:
\begin{align}
    A &\approx \sum_{i} \frac{1}{2} [g(\theta_i)]^2 \Delta \theta_i \nonumber \\
    \intertext{Taking a Riemann sum:}
    &= \int_{\alpha}^{\beta} \frac{1}{2} [g(\theta)]^2 \, d\theta
\end{align}
\begin{example}
    Given $r = 1 + \cos \theta$, we have:
    \begin{align*}
        A &= \int_0^{2\pi} \frac{1}{2} (1 + \cos \theta)^2 \, d\theta \\
        &= \int_0^{2\pi} \frac{1}{2} (1 + 2\cos \theta + \cos^2 \theta) \, d\theta \\
        &= \frac{1}{2}\int_0^{2\pi} 2 + 2\cos \theta + \frac{1 + \cos 2\theta}{2} \, d\theta \\
        &= \frac{1}{2}\left(2\pi + 2\sin \theta + \frac{\theta}{2} + \frac{\sin 2\theta}{4}\right) \Big|_0^{2\pi}= \frac{3\pi}{2}
    \end{align*}
\end{example}
\paragraph{Area between Two Polar Curves} The area between two polar curves $r^2 = f(\theta)$ and $r = g(\theta)$ is given by:
\begin{equation}
    A = \int_{\alpha}^{\beta} \frac{1}{2} [f(\theta)^2 - g(\theta)^2] \, d\theta
\end{equation}
\begin{example}
    Given lemniscate $r^2 = 4\cos 2\theta$, circle $r = 1$, to find the area between the two curves, outside the circle, and inside the lemniscate. We have:
    \begin{align*}
        \intertext{We first consider the intersection points}
        4\cos 2\theta &= 1 \\
        \cos 2\theta &= \frac{1}{4} \\
        \intertext{We have points at $\theta = \pm 0.659$. Also, since the lemniscate is symmetric about the $y$-axis, we have:}
        \frac{1}{2}A &= \int_{-0.659}^{0.659} 4\cos^2 2\theta - 1 \, d\theta \\
        A &= 2.554
    \end{align*}
\end{example}
\begin{example}
    Given $r = \sin \theta$, $r = \cos \theta$, we have:
    \begin{align*}
        \intertext{We first consider the intersection points}
        \sin \theta &= \cos \theta \\
        \tan \theta &= 1 \\
        \theta &= \frac{\pi}{4}
        \intertext{We then consider the area between the two curves, notice that (0, 0) is also an intersection point.}
        A &= \frac{1}{2}\int_0^{\frac{\pi}{4}}  (\sin \theta)^2 - (\cos \theta)^2 \, d\theta \\
        &= \frac{\pi - 2}{8}
    \end{align*}
\end{example}
\paragraph{Tangent Lines in Polar Coordinates} The tangent line to a polar curve $r = r(\theta)$ is derived as follows:
\begin{align}
    x(\theta) &= r(\theta)\cos \theta, \quad y(\theta) = r(\theta)\sin \theta \nonumber \\
    \intertext{By chain rule, we have and product rule, we have:}
    \frac{dy}{dx} &= \frac{r'(\theta)\sin \theta + r(\theta)\cos \theta}{r'(\theta)\cos \theta - r(\theta)\sin \theta} \\
\end{align}
\paragraph{Length of a Polar Curve} Derived from above, the length of a polar curve $r = r(\theta)$ between $\alpha$ and $\beta$ is given by:
\begin{equation}
    L = \int_{\alpha}^{\beta} \sqrt{[r(\theta)]^2 + [r'(\theta)]^2} \, d\theta
\end{equation}
\begin{example}
    Given $r = a - a\cos \theta$, we have:
    \begin{align*}
        r' &= a\sin \theta \\
        L &= \int_0^{2\pi} \sqrt{(a - a\cos \theta)^2 + a^2\sin^2 \theta} \, d\theta \\
        &= a\int_0^{2\pi} \sqrt{2 - 2\cos \theta} \, d\theta \\
        \intertext{We can use the double angle formula to simplify the integral, with $\cos 2\theta = 1 - 2\sin^2 \theta$:}
        &= 8a
    \end{align*}
\end{example}
\section{Infinite Sequences and Series}
\subsection{Sequences}
\paragraph{Sequences} A sequence is a list of numbers written in a specific order.
\begin{example}
    Let $f: A \to \mathbb{R}$ be a function, and $f(x) = \frac{1}{x}$. We can restrict the domain of $f$ to $A = \{1, 2, 3, \ldots\} = \mathbb{N}$, and we can define the sequence $a_n = f(n) = \frac{1}{n}$.
\end{example}
\paragraph{Notation} Typcially, we denote a sequence as $\{a_n\}$, where $a_n$ is the $n$th term of the sequence. We can also denote with curly bracket as $\{a_n\} = \{a_1, a_2, a_3, \ldots\} = \{f(n)\}$.
\paragraph{Ratio} The ratio of two sequences $\{a_n\}$ and $\{b_n\}$ for some $n \le N$ is a sequence $\{c_n\}$ such that $c_n = \frac{a_n}{b_n}$.
\begin{example}
    Let $a_n = n^2$, $b_n = e^n$, we have:
    $$\frac{a_n}{b_n} = \{\frac{1}{e}, \frac{4}{e^2}, \frac{9}{e^3}, \ldots\}$$
\end{example}
\paragraph{Bounded Above} A sequence $\{a_n\}$ is said to be bounded above if there exists a number $M$ such that $|a_n| \le M$ for all $n$. (Similarly for bounded below).
\begin{definition}[Monotonic Sequences]
    A sequence $\{a_n\}$ is said to be non decreasing if $a_{n+1} \ge a_n$ for all $n$. Similarly, a sequence is said to be non increasing if $a_{n+1} \le a_n$ for all $n$. A sequence is said to be increasing if $a_{n+1} > a_n$ for all $n$. Similarly, a sequence is said to be decreasing if $a_{n+1} < a_n$ for all $n$.
\\ A sequence $\{a_n\}$ is said to be monotonic if it is either increasing, decreasing, non increasing, or non decreasing.
\end{definition}
\begin{definition}[Unbounded Sequences]
    A sequence $\{a_n\}$ is said to be unbounded if it is not bounded above or below. We can prove it using a epsilon-delta proof or considering the derivatives of the function that defines the sequence.
\end{definition}
\begin{example}
    Proof that the sequence $\{a_n\} = 2^n$ is unbounded. Let $M$ be a number, we can find an $n$ such that $2^n > M$. We can find this $n$ by taking the logarithm of $M$ to the base $2$.
\end{example}
\begin{example}
    Proof that the sequence $\{a_n\} = \frac{x^2}{e^x}$ is bounded above. We show that the derivative is negative after a certain point ($x = 2$), thus the sequence is bounded above.
\end{example}
\subsection{The Limit of a Sequence}
\begin{definition}[Limit of a Sequence]
    The limit of a sequence $\{a_n\}$ is given by:
    \begin{equation*}
        \lim_{n \to \infty} a_n = L
    \end{equation*}
    if for every $\epsilon > 0$, there exists an $N$ such that $|a_n - L| < \epsilon$ for all $n > N$. We denote this as $a_n \to L$ as $n \to \infty$.
\end{definition}
\paragraph{Convergence} A sequence $\{a_n\}$ is said to converge to $L$ if the limit of the sequence is $L$. And the following are true:
\begin{enumerate}
    \item If a sequence converges, then it is bounded.
    \item If a sequence is unbounded, then it does not converge, i.e. it diverges.
    \item However, if a sequence is bounded, it may or may not converge.
\end{enumerate}
\begin{example}
    Prove that the sequence $\{a_n\} = \frac{n}{n+1}$ converges to $1$. We can show that $|a_n - 1| < \epsilon$ for all $n > N$ by taking $N = \frac{1}{\epsilon}$ by some algebraic manipulation. We have:\\
    Given $\epsilon > 0$, let $N = \frac{1}{\epsilon}$, then for all $n > N$:
    \begin{align*}
        |a_n - 1| &= |\frac{n}{n+1} - 1| \\
        &= |\frac{n - n - 1}{n+1}| \\
        &= \frac{1}{n+1} < \frac{1}{n} < \epsilon
    \end{align*}
\end{example}
\begin{theorem}[Uniqueness of Limits]
    The limit of a sequence is unique. That is, of $\lim_{n \to \infty} a_n = L$ and $\lim_{n \to \infty} a_n = M$, then necessarily $L = M$.
\end{theorem}
\begin{theorem}[Monotone Convergence Theorem]
    A monotonic sequence is convergent if and only if it is bounded. And necessarily the following are true if the sequence converges:
    \begin{enumerate}
        \item If the sequence is increasing or non decreasing, then the limit is the supremum of the sequence.
        \item If the sequence is decreasing or non increasing, then the limit is the infimum of the sequence.
    \end{enumerate}
\end{theorem}
\begin{theorem}[Limit Operations]
    Given two sequences $\{a_n\}$ and $\{b_n\}$, and $c \in \mathbb{R}$, we have:
    \begin{enumerate}
        \item \textbf{Sum of Limits}: $\lim_{n \to \infty} (a_n + b_n) = \lim_{n \to \infty} a_n + \lim_{n \to \infty} b_n$
        \item \textbf{Product of Limits}: $\lim_{n \to \infty} (a_n \cdot b_n) = \lim_{n \to \infty} a_n \cdot \lim_{n \to \infty} b_n$
        \item \textbf{Scalar Multiple of Limits}: $\lim_{n \to \infty} c \cdot a_n = c \cdot \lim_{n \to \infty} a_n$
        \item \textbf{reciprocal of Limits}: $\lim_{n \to \infty} \frac{1}{a_n} = \frac{1}{\lim_{n \to \infty} a_n}$
        \item \textbf{Quotient of Limits}: $\lim_{n \to \infty} \frac{a_n}{b_n} = \frac{\lim_{n \to \infty} a_n}{\lim_{n \to \infty} b_n}$, provided $\lim_{n \to \infty} b_n \neq 0$
    \end{enumerate}
\end{theorem}
\begin{theorem}[Squeeze Theorem\footnote{Also known as the Sandwich Theorem or the Pinching Theorem}]
    Given three sequences $\{a_n\}$, $\{b_n\}$, and $\{c_n\}$, if $a_n \le b_n \le c_n$ for all $n$, and $\lim_{n \to \infty} a_n = \lim_{n \to \infty} c_n = L$, then necessarily $\lim_{n \to \infty} b_n = L$. This is useful for proving limits of sequences.
\end{theorem}
\begin{theorem}[Composite Function]
    Given sequence $\{a_n\}$ covnerges to $L$ and continuous function $f$ at $L$, we have:
    \begin{equation}
        \lim_{n \to \infty} f(a_n) = f(\lim_{n \to \infty} a_n)
    \end{equation}
\end{theorem}
\paragraph{Some Common Limits} Some common limits of sequences are:
\begin{enumerate}
    \item for $a > 0$, $\lim_{n \to \infty} a^{1/n} = 1$
    \item for $|a| < 1$, $\lim_{n \to \infty} a^n = 0$
    \item for $a > 0$, $\lim_{n \to \infty} \frac{1}{n^a} = 0$
    \item for $a > 0$, $\lim_{n \to \infty} \frac{a^n}{n!} = 0$
\end{enumerate}
\subsection{Series}
\begin{definition}[Infinite Series]
    An infinite series is the sum of the terms of an infinite sequence. We denote the sum of the series as:
    \begin{equation}
        S = \sum_{n=1}^{\infty} a_n = a_1 + a_2 + a_3 + \ldots
    \end{equation}
    We also denote it as just $\sum a_n$.
\end{definition}
\begin{definition}[Partial Sum]
    The partial sum of an infinite series is the sum of the first $n$ terms of the series. We denote the partial sum as:
    \begin{equation}
        S_n = \sum_{i=1}^{n} a_i = a_1 + a_2 + \ldots + a_n
    \end{equation}
    If the sequance $\{S_n\}$ converges to $S$, then the series is said to converge to $S$; otherwise, it is said to diverge.
\end{definition}
\paragraph{Geometric Series} A geometric series is a series of the form:
\begin{equation}
    \sum_{n=0}^{\infty} ar^n = a + ar + ar^2 + \ldots
\end{equation}
\begin{theorem}[Geometric Series]
    A geometric series converges if $|r| < 1$, and the sum of the series is given by:
    \begin{equation}
        S = \frac{a}{1 - r}
    \end{equation}
    If $|r| \ge 1$, then the series diverges.
\end{theorem}
\begin{theorem}[Addition of Series]
    Given two series $\sum a_n$ and $\sum b_n$, we have:
    \begin{equation}
        \sum a_n + \sum b_n = \sum (a_n + b_n)
    \end{equation}
\end{theorem}
\begin{theorem}[Convergence]
    $\sum a_n$ converges if and only if $\sum_{k=j}^{\infty} a_k$ converges for all $j$.
\end{theorem}
\begin{example}
    Given $\sum \frac{3^{k-1}}{4^{3k+1}}$ converges. Then we have:
    \begin{align*}
        \sum \frac{3^{k-1}}{4^{3k+1}} = \frac{1}{12} + \frac{1}{64} + \frac{1}{4^7} + \ldots_n + \sum^{\infty}_{k=n} \frac{3^{k-1}}{4^{3k+1}}
    \end{align*}
    converges for all $n$.
\end{example}
\begin{theorem}[Zero Test, Divergence Test]
    If $\lim_{n \to \infty} a_n \ne 0$, then $\sum a_n$ diverges. Note that the contrapositive is true, i.e. if $\sum a_n$ converges, then $\lim_{n \to \infty} a_n = 0$.
\end{theorem}
\subsection{The Integral Test and Estimates of Sums}
\begin{theorem}[Integral Test]
    Let $f$ be a continuous, positive, and decreasing function on $[1, \infty)$, and let $a_n = f(n)$. Then the series $\sum a_n$ converges if and only if the improper integral $\int_1^{\infty} f(x) \, dx$ converges.
\end{theorem}
\begin{example}[$p$=2-series]
    Given $\sum \frac{1}{n^2}$, we can use the integral test to show that the series converges. We have:
    \begin{align*}
        \int_1^{\infty} \frac{1}{x^2} \, dx &= -\frac{1}{x} \Big|_1^{\infty} \\
        &= 1
    \end{align*}
    Thus, the series converges.
\end{example}
Note that the ``iff'' statement allow us to determine the convergence of an integral.
\begin{example}[Harmonic Series]
    Given $\sum \frac{1}{n}$, we can use the integral test to show that the series diverges. We have:
    \begin{align*}
        \int_1^{\infty} \frac{1}{x} \, dx &= \ln x \Big|_1^{\infty} \\
        &= \infty
    \end{align*}
    Thus, the series diverges. The harmonic series is a classic example of a divergent series.
\end{example}
\begin{definition}[Remainder]
    The remainder of a series $\sum a_n$ is given by:
    \begin{equation}
        R_n = S - S_n = \sum_{k=n+1}^{\infty} a_k
    \end{equation}
    We can use the remainder to estimate the sum of a series.
\end{definition}
\paragraph{Estimating Sums} We can estimate the sum of a series by using the remainder of the series. We can use the remainder to estimate the sum of a series. Note that we always have:
\begin{equation}
    R_n \le \int_n^{\infty} f(x) \, dx /quad \text{and} \quad R_n \ge \int_{n+1}^{\infty} f(x) \, dx
\end{equation}
We can use it to estimate the sum of a series from a partial sum.
\begin{equation}
    S_n + \int_{n+1}^{\infty} f(x) \, dx \le S \le S_n + \int_n^{\infty} f(x) \, dx
\end{equation}
\begin{example}
    Given $\sum \frac{1}{n^2}$, we can use the remainder to estimate the sum of the series. We have:
    \begin{align*}
        S_n + \int_{n+1}^{\infty} \frac{1}{x^2} \, dx &\le S \le S_n + \int_n^{\infty} \frac{1}{x^2} \, dx \\
        \frac{1}{n} + \int_{n+1}^{\infty} \frac{1}{x^2} \, dx &\le S \le \frac{1}{n} + \int_n^{\infty} \frac{1}{x^2} \, dx
    \end{align*}
    We can then evaluate the integrals to obtain the value bounds.
\end{example}
\subsection{Comparison Tests}
\begin{theorem}[Basic Comparison Test]
    Let $\{a_n\}$ and $\{b_n\}$ be two sequences such that $0 \le a_n \le b_n$ for all $n$. Then:
    \begin{enumerate}
        \item If $\sum b_n$ converges, then $\sum a_n$ converges.
        \item If $\sum a_n$ diverges, then $\sum b_n$ diverges.
    \end{enumerate}
\end{theorem}
\begin{proof}
    Assume, WLOG, that $a_k \le b_k$ for all $k$. Define $S_n = \sum_{k=1}^{n} a_k$ and $T_n = \sum_{k=1}^{n} b_k$. Also, $T = \sum b_k$ exists. We have:
    \begin{align*}
        S_{n+1} &= S_n + a_{n+1} > S_n
        \intertext{Thus, $\{S_n\}$ is increasing. Also, we have:}
        S_n &\le T_n < T
        \intertext{Thus, $\{S_n\}$ is bounded above. Thus, $\{S_n\}$ converges by the Monotone Convergence Theorem.}
    \end{align*}
\end{proof}
\begin{example}
    Let $S = \sum \frac{7}{17n^2 + 3n^{1/3} + 5}$ We know that $17n^2 > 17n^{1/3} + 5$ for all $n$. Thus, we have:
    \begin{align*}
        \frac{7}{17n^2} &> \frac{7}{17n^2 + 3n^{1/3} + 5} \\
        \intertext{We can then use the comparison test to show that the series converges.}
    \end{align*}
\end{example}
\appendix
\input{appendix.tex}
\end{document}