### 2025-09-01 YEAR2/CHE260_LectureNotes/main.tex
pre year2

\documentclass[11pt]{report}
\input{preamble}
\begin{document}
\thispagestyle{empty}
{\LARGE \bf CHE 260 Lecture Notes}\\
{\large Hei Shing Cheung}\\
Thermodynamics and Heat Transfer, Fall 2025 \hfill CHE260\\
\\
The up-to-date version of this document can be found at \url{https://github.com/HaysonC/skulenotes}\\

\chapter{Thermodynamics}

\section{Systems and Properties in Thermodynamics}

\chapter{Heat Transfer}

\end{document}