### 2025-09-16 YEAR2/PHY293_LectureNotes/main.tex

\documentclass[11pt]{report}
\input{preamble}
\begin{document}
\thispagestyle{empty}
{\LARGE \bf PHY 293 Lecture Notes}\\
{\large Hei Shing Cheung}\\
Waves and Modern Physics, Fall 2025 \hfill PHY293\\
\\
The up-to-date version of this document can be found at \url{https://github.com/HaysonC/skulenotes}\\

\chapter{Waves}
\section{Harmonic Oscillators}

\subsection{Governing Equations of Harmonic Oscillators}
This subsection collects the baseline ODEs for simple, damped, and driven oscillators to set notation used later.

\paragraph{Types of Harmonic Oscillators} There are three types of harmonic oscillators: simple, damped, and driven harmonic oscillators. Consider a simple one dimensional harmonic oscillator, they are defined by the following differential equations:

\begin{definition}[Simple Harmonic Oscillator]
    A simple harmonic oscillator is described by Hooke's law:

    \begin{equation} \label{eq:hooke}
        m \frac{d^2 x}{dt^2} + kx = 0
    \end{equation}
    where \( k \) is the spring constant, \( m \) is the mass, and \( x \) is the displacement from equilibrium.
\end{definition}

\begin{definition}[Damped Harmonic Oscillator]
    A damped harmonic oscillator is described by the following differential equation, by adding a damping term proportional to $\dot{x}$ to the simple harmonic oscillator equation:
    \begin{equation} \label{eq:damped_ho}
        m \frac{d^2 x}{dt^2} + b \frac{dx}{dt} + kx = 0   
    \end{equation}
    where \( b \) is the damping coefficient.
\end{definition}

\begin{shaded}
	\textbf{Note on damping parameter conventions.} Different texts use different symbols and normalizations:
\begin{itemize}
    \item Our notes use $\gamma = \tfrac{b}{m}$, so the ODE reads $\ddot{x} + \gamma\dot{x} + \omega_0^2 x = 0$ and the decay envelope is $e^{-\gamma t/2}$.
    \item Many texts define $\beta = \tfrac{b}{2m}$ and may call it $\gamma$ instead. In that convention the ODE is $\ddot{x} + 2\beta\dot{x} + \omega_0^2 x = 0$ with envelope $e^{-\beta t}$.
\end{itemize}
Equivalences: $\beta = \tfrac{\gamma}{2}$ and our underdamped condition $\gamma < 2\omega_0$ corresponds to $\beta < \omega_0$ in the $2\beta$ convention. When comparing formulas, check which definition is being used and replace $\gamma \leftrightarrow 2\beta$ accordingly.
\end{shaded}

\begin{definition}[Driven Harmonic Oscillator]
    A driven harmonic oscillator is described by the following differential equation, which includes an external driving force $F(t)$:
    \begin{equation}
        m \frac{d^2 x}{dt^2} + b \frac{dx}{dt} + kx = F(t)   
    \end{equation}
\end{definition}
% [Moved: The Wave Equation section is presented later for a smoother flow after oscillators]

\subsection{Simple Harmonic Motion}
We analyze the undamped solution forms, relate constants to initial conditions, and derive period/frequency relations.
\begin{definition}[Simple Harmonic Motion]
    You should have learned Hooke's law and Newton's second law, which give us the equation of motion for a simple harmonic oscillator. The same as equation (\ref{eq:hooke}), which can be rewritten as:
    \begin{equation}
        F = m\ddot{x} = -kx
    \end{equation}
    By setting $\omega^2 = \frac{k}{m}$, a general solution can be written as:
    \begin{equation}
        x(t) = x_0 + A_1 \cos(\omega t) + A_2 \sin(\omega t)
    \end{equation}
    where $A_1$ and $A_2$ are constants determined by the IVP, $\omega$ is the angular frequency, and $\phi$ is the phase constant. $x_0$ is the equilibrium position (often set to 0). The unknown constants can be determined by knowing $x, \dot{x}$ at specific times.
\end{definition}

\begin{definition}[Period, Frequency, and Angular Frequency]
    The period \( T \) is the time it takes for one complete cycle of the motion, given by:
    \begin{equation}
        T = 2\pi \sqrt{\frac{m}{k}}
    \end{equation}
    The frequency \( f \) is the number of cycles per unit time, given by:
    \begin{equation}
        f = \frac{1}{T} = \frac{1}{2\pi} \sqrt{\frac{k}{m}}
    \end{equation}
    The angular frequency \( \omega \) is related to the frequency by:
    \begin{equation}
        \omega = 2\pi f = \sqrt{\frac{k}{m}}
    \end{equation}

\end{definition}

\begin{example}
    A simple harmonic oscillator consisting of mass \(m = 11.0\ \mathrm{kg}\) attached to a spring with spring constant \(k = 201\ \mathrm{N\,m^{-1}}\). At time \(t=0\ \mathrm{s}\) the oscillator is at position \(x(0) = -0.207\ \mathrm{m}\) and has velocity \(v(0) = -1.33\ \mathrm{m\,s^{-1}}\). Determine all coefficients of the equation describing the position \(x(t)\) of the oscillator as a function of time, assuming the offset is zero.

    To solve for $A_1$ and $A_2$, while we assume $x_0 = 0$, we can use the initial conditions:
    \begin{align*}
        x(0) &= A_1 \cos(0) + A_2 \sin(0) = A_1 = -0.207\ \mathrm{m} \\
        v(0) &= -A_1 \omega \sin(0) + A_2 \omega \cos(0) = A_2 \omega = -1.33\ \mathrm{m\,s^{-1}}
    \end{align*}

    We can find $\omega$ from the given $m$ and $k$:
    $$
    \omega = \sqrt{\frac{k}{m}} = \sqrt{\frac{201\ \mathrm{N\,m^{-1}}}{11.0\ \mathrm{kg}}} \approx 4.28\ \mathrm{rad\,s^{-1}}
    $$
    Therefore, we can solve for $A_2$:
    $$
    A_2 = \frac{v(0)}{\omega} = \frac{-1.33\ \mathrm{m\,s^{-1}}}{4.28\ \mathrm{rad\,s^{-1}}} \approx -0.311\ \mathrm{m}
    $$
    Thus, the equation describing the position \(x(t)\) of the oscillator as a function of time is:
    $$
    x(t) = -0.207 \cos(4.28 t) - 0.311 \sin(4.28 t)
    $$
\end{example}
\begin{theorem}[A Trigonometric Identity]
We can also express the solution in a more compact form using a single cosine function with a phase shift:
\begin{equation}
    x(t) = A \cos(\omega t + \phi)
\end{equation}
where
\begin{subequations}
    \begin{align}
        A &= \sqrt{A_1^2 + A_2^2}, \\
        \phi &= \arctan\!\left(\frac{-A_2}{A_1}\right) 
              = \arctan\!\left(\frac{-v(0)/\omega}{x(0)}\right).
    \end{align}
\end{subequations}
\end{theorem}

\begin{proof}
Let $A = \sqrt{A_1^2 + A_2^2}$ and choose $\phi$ such that 
\[
\cos(\phi) = \frac{A_1}{A}, \qquad \sin(\phi) = -\frac{A_2}{A}.
\]
Then, we can rewrite our original solution as
\begin{align*}
    x(t) &= A_1 \cos(\omega t) + A_2 \sin(\omega t) \\
         &= A \cos(\phi) \cos(\omega t) - A \sin(\phi) \sin(\omega t) \\
         &= A \big[\cos(\phi)\cos(\omega t) - \sin(\phi)\sin(\omega t)\big] \\ 
         &= A \cos(\omega t + \phi),
\end{align*}
by the cosine addition formula.
\end{proof}

\begin{example}
    To determine the amplitude \(A\) and phase constant \(\phi\) for the oscillator in the previous example, we can use the values of \(A_1\) and \(A_2\) we found:
    \begin{align*}
        A &= \sqrt{(-0.207)^2 + (-0.311)^2} \approx 0.374\ \mathrm{m} \\
        \phi &= \arctan\!\left(\frac{-(-0.311)}{-0.207}\right) \approx 4.12\ \mathrm{rad} \quad (\text{since } A_1 < 0 \text{ and } A_2 < 0)
    \end{align*}
    Therefore, the equation describing the position \(x(t)\) of the oscillator as a function of time can also be written as:
    $$
    x(t) = 0.374 \cos(4.28 t + 4.12)
    $$
\end{example}

\begin{definition}[The Energy of a Simple Harmonic Oscillator]
    The total mechanical energy \(E\) of a simple harmonic oscillator is the sum of its kinetic energy \(K\) and potential energy \(U\). 
    \begin{equation}
        E = K + U
    \end{equation}
    First we consider the change of potential energy from a position $x_i$ to $x_f$, assuming the path is along the spring or the curve $C$ of the oscillator. The force exerted by the spring is given by Hooke's law, \( F = -kx \). The change in potential energy can be simply parametized and calculated as follows:
    \begin{equation}
        \Delta U = \int_C F\cdot ds = -\int_{x_i}^{x_f} F\,dx = \int_{x_i}^{x_f} kx\,dx = \left[\frac{1}{2}kx^2\right]_{x_i}^{x_f} = \frac{1}{2}k(x_f^2 - x_i^2)
    \end{equation}

    Therefore, the potential energy \(U\) at a position \(x\) (taking the reference point at \(x=0\)) is given by:
    \begin{equation}
        U(x) = \frac{1}{2}kx^2
    \end{equation}

    The kinetic energy \(K\) of the oscillator is given by:
    \begin{equation}
        K = \frac{1}{2}m\dot{x}^2
    \end{equation}
    Therefore, the total mechanical energy \(E\) of the simple harmonic oscillator is:
    \begin{equation}
        E = K + U = \frac{1}{2}m\dot{x}^2 + \frac{1}{2}kx^2
    \end{equation}
    The total mechanical energy \(E\) remains constant over time, as energy is conserved in the absence of non-conservative forces (like friction or air resistance).

\end{definition}

\subsection{Damped Harmonic Motion}
We solve the damped ODE, classify regimes (underdamped, critical, overdamped), and connect decay rates with parameters.
\begin{definition}[Damped Harmonic Motion]
    For small velocities, the drag force is approximately proportional to the velocity and acts in the opposite direction. This drag force can be modeled as \( F_d = -\gamma \dot{x} \), where \( \gamma \) is the damping coefficient. Including this drag force in the equation of motion for a harmonic oscillator leads to the damped harmonic oscillator equation (\ref{eq:damped_ho}). Which could be rewritten as:
    \begin{equation}
        \ddot{x} + \gamma \dot{x} + \omega_0^2 x = 0
    \end{equation}
    where \( \omega_0 = \sqrt{\frac{k}{m}} \) is the natural angular frequency of the undamped oscillator, and \( \gamma = \frac{b}{m} \) is the damping coefficient per unit mass.

    To skip the math, lets assume a solution of the form \( x(t) = e^{i\omega t} \), substituting into the differential equation gives us a formulation for \( \omega \):
    \begin{align}
        -\omega^2 - i\gamma\omega + \omega_0^2 &= 0 \nonumber \\
        \omega &= -i\frac{\gamma}{2} \pm \sqrt{\omega_0^2 - \frac{\gamma^2}{4}} \label{eq:omega_damped}
    \end{align}
    Also, we can characterize the real and imaginary parts of \( \omega \) as:
    \begin{subequations}
        \begin{equation}
            \omega_r = \mathrm{Re}(\omega) = \pm \sqrt{\omega_0^2 - \frac{\gamma^2}{4}}
        \end{equation}
        \begin{equation}
            \omega_i = \mathrm{Im}(\omega) = -\frac{\gamma}{2}
        \end{equation}
    \end{subequations}
    The general solution for the damped harmonic oscillator can be written as:
    \begin{equation}
        x(t) = \exp(\omega_i t) \exp(-i \omega_r t) = \exp\left(-\frac{\gamma}{2}t\right) \exp(\mp i \sqrt{\omega_0^2 - \frac{\gamma^2}{4}} t)
    \end{equation}
    \begin{subequations}
    \begin{itemize}
        \item \textbf{No Damping} (\( \gamma = 0 \)): The system behaves like a simple harmonic oscillator with angular frequency \( \omega_0 \). Given by:
        \begin{equation}
            z = \exp(-i\omega_0 t)
        \end{equation}
        \item \textbf{Underdamping} (\( 0 < \gamma < 2\omega_0 \)): The system oscillates with a gradually decreasing amplitude. The angular frequency of oscillation is given by \( \omega_r = \sqrt{\omega_0^2 - \frac{\gamma^2}{4}} \). Given by:
        \begin{equation}
            z = \exp\left(-\frac{\gamma}{2}t\right) \exp(-i \omega_r t)
        \end{equation}
        The trigonometric form of the solution is:
        \begin{equation}
            x(t) = A_0 \exp\left(-\frac{\gamma}{2}t\right) \cos(\omega_r t + \phi)
        \end{equation}
        where \( A_0 \) and \( \phi \) are constants determined by the initial conditions
        From this, we can derive the following cases:
        \item \textbf{Critical Damping} (\( \gamma = 2\omega_0 \)): The system returns to equilibrium as quickly as possible without oscillating. Consider:
        $$
            x(t) = e^{-\frac{\gamma}{2}t}f(t)
        $$
        Inserting into the differential equation, we get:
        $$
            \ddot{f} + \left(\omega_0^2 - \frac{\gamma^2}{4}\right)f = 0
        $$
        Since \( \gamma = 2\omega_0 \), we have \( \omega_0^2 - \frac{\gamma^2}{4} = 0 \), leading to:
        $$
            \ddot{f} = 0 \implies f(t) = A_1 t + A_2
        $$
        Therefore, the general solution for the critically damped case is:
        \begin{equation}
            x(t) = (A_1 t + A_2) \exp\left(-\frac{\gamma}{2}t\right)
        \end{equation}
        where \( A_1 \) and \( A_2 \) are constants determined by
        \item \textbf{Overdamping} (\( \gamma > 2\omega_0 \)): The system returns to equilibrium without oscillating, but more slowly than in the critically damped case. The solution is given by:
        \begin{equation}
            z = \exp\left(-\frac{\gamma}{2}t\right) \exp\left(\sqrt{\frac{\gamma^2}{4} - \omega_0^2} t\right)
        \end{equation}
        So the general solution is (the solution is via a substitution of $x(t) = e^{-\gamma t/2} f(t)$ into the differential equation, which resolves the ODE to a simple form):
        \begin{equation}
            x(t) = A_1 \exp\left[\left(-\frac{\gamma}{2} + \sqrt{\frac{\gamma^2}{4} - \omega_0^2}\right) t\right] + A_2 \exp\left[\left(-\frac{\gamma}{2} - \sqrt{\frac{\gamma^2}{4} - \omega_0^2}\right) t\right]
        \end{equation}
        where \( A_1 \) and \( A_2 \) are constants determined by the initial conditions.
    \end{itemize}
    \end{subequations}
\end{definition}

\subsection{Energy and Quality Factor}
We study how energy decays under light damping, define the time constant and quality factor Q, and relate them to response.
\begin{definition}[Energy of a Very Light Damping]
    Consider a very lightly damped harmonic oscillator, where \( \gamma \ll \omega_0 \). In this case, the angular frequency of oscillation \( \omega_r \) can be approximated as:
    $$
    \omega_r \approx \omega_0 \left(1 - \frac{\gamma^2}{8\omega_0^2}\right) \approx \omega_0
    $$
    So the motion of the lightly damped oscillator can be approximated as:
    $$
    x(t) \approx A_0 e^{-\frac{\gamma}{2}t} \cos(\omega_0 t + \phi)
    $$
    Then, we can calculate the velocity of the oscillator:
    \begin{align*}
        \dot{x}(t) &= -\frac{\gamma}{2} A_0 e^{-\frac{\gamma}{2}t} \cos(\omega_0 t + \phi) - A_0 \omega_0 e^{-\frac{\gamma}{2}t} \sin(\omega_0 t + \phi) \\
        &= A_0 \omega_0 e^{-\frac{\gamma}{2}t} \left(-\frac{\gamma}{2\omega_0} \cos(\omega_0 t + \phi) - \sin(\omega_0 t + \phi)\right)
    \end{align*}
    The total mechanical energy \( E(t) \) of the lightly damped oscillator is given by:
    \begin{align*}
        E(t) &= \frac{1}{2}m\dot{x}^2 + \frac{1}{2}kx^2 \\
             &= \frac{1}{2}m \left[A_0 \omega_0 e^{-\frac{\gamma}{2}t} \left(-\frac{\gamma}{2\omega_0} \cos(\omega_0 t + \phi) - \sin(\omega_0 t + \phi)\right)\right]^2 + \frac{1}{2}k \left[A_0 e^{-\frac{\gamma}{2}t} \cos(\omega_0 t + \phi)\right]^2 \\
             &= \frac{1}{2}m A_0^2 \omega_0^2 e^{-\gamma t} \left[\left(-\frac{\gamma}{2\omega_0} \cos(\omega_0 t + \phi) - \sin(\omega_0 t + \phi)\right)^2 + \cos^2(\omega_0 t + \phi)\right] \\
             &= \frac{1}{2}m A_0^2 \omega_0^2 e^{-\gamma t} \left[\sin^2(\omega_0 t + \phi) + \cos^2(\omega_0 t + \phi) + \frac{\gamma^2}{4\omega_0^2} \cos^2(\omega_0 t + \phi) + \frac{\gamma}{\omega_0} \sin(\omega_0 t + \phi) \cos(\omega_0 t + \phi)\right] \\
             &\approx \frac{1}{2}m A_0^2 \omega_0^2 e^{-\gamma t} \left[1 + \frac{\gamma^2}{4\omega_0^2} \cos^2(\omega_0 t + \phi)\right] \quad (\text{neglecting the small term } \frac{\gamma}{\omega_0} \sin(\omega_0 t + \phi) \cos(\omega_0 t + \phi)) \\
             &\approx \frac{1}{2}m A_0^2 \omega_0^2 e^{-\gamma t} \quad (\text{since } \frac{\gamma^2}{4\omega_0^2} \text{ is very small}) \\
             &= E_0 e^{-\gamma t} \quad \text{where } E_0 = \frac{1}{2}m A_0^2 \omega_0^2 \text{ is the initial energy at } t=0
    \end{align*}
    We can also define the time constant \( \tau \) as the time it takes for the energy to decrease to \( \frac{1}{e} \) of its initial value:
    \begin{equation}
        \tau = \frac{1}{\gamma}
    \end{equation}
    So we have, for very light damping:
    \begin{equation}
        E(t) = E_0 e^{-\gamma t} = E_0 e^{-\frac{t}{\tau}}
    \end{equation}
\end{definition}

\begin{definition}[Rate of Energy Loss]
    Taking the time derivative of the total mechanical energy \( E(t) \):
    \begin{align*}
        \frac{dE}{dt} &= \frac{d}{dt} \left(\frac{1}{2}m\dot{x}^2 + \frac{1}{2}kx^2 \right) \\
                      &= (ma + kx)\dot{x} \\
    \end{align*}
    For a undamped harmonic oscillator, \( ma + kx = 0 \), so \( \frac{dE}{dt} = 0 \), indicating that the total mechanical energy is conserved and obeys Hooke's law completely. However, for a damped harmonic oscillator, \( ma + kx = -b\dot{x} \), leading to:
    \begin{equation}
        \frac{dE}{dt} = -b\dot{x}^2
    \end{equation}
\end{definition}

\begin{definition}[Quality Factor (Q-Factor)]
    The quality factor \( Q \) is a dimensionless parameter that characterizes the damping of a harmonic oscillator. It is defined as:
    \begin{subequations}
    \begin{equation}
        Q = \frac{\omega}{\gamma} = \omega \tau
    \end{equation}
    And for very light damping, we can approximate \( \omega \approx \omega_0 \), leading to:
    \begin{equation}
        Q \approx \frac{\omega_0}{\gamma} = \omega_0 \tau
    \end{equation}
    \end{subequations}
    This allows us to rewrite the equation of a damped harmonic oscillator as:
    \begin{equation}
        \ddot{x} + \frac{\omega_0}{Q} \dot{x} + \omega_0^2 x = 0
    \end{equation}
    and:
    \begin{equation}
        \omega = \omega_0 \sqrt{1 - \frac{1}{4Q^2}}
    \end{equation}

We can also consider the ratio between the energy at one time and the energy one period later:
\begin{align*}
    \frac{E(t + T)}{E(t)} &= \frac{E_0 e^{-\gamma (t + T)}}{E_0 e^{-\gamma t}} \\
    &= e^{-\gamma T} \approx 1 - \gamma T \\
    \frac{E(t + T) - E(t)}{E(t)} & \approx -\gamma T = -\frac{2\pi}{\omega_0} \gamma = -\frac{2\pi}{Q}
\end{align*}
\end{definition}

\begin{example}
    \textit{What is the number of radians through which the damped system oscillates as its energy decreases to $1/e$ of its initial value?}

    We have
    $$
    \frac{E}{E_0} = e^{-\gamma t} = \frac{1}{e} \implies \gamma t = 1 \implies t = \frac{1}{\gamma} = \tau
    $$
    So the number of radians is
    $$
    \theta = \omega \tau = \frac{\omega}{\gamma} = Q
    $$

\end{example}

\subsection{Undamped Forced Oscillations}
We examine steady-state response to sinusoidal driving without losses and identify resonance behavior.
\begin{definition}[Undamped Forced Oscillations]
    Consider a driver force acting on the mass-spring system; the ODE becomes:
    \begin{equation}
        m\ddot{x} + kx = \eta
    \end{equation}
    where \( \eta = F_0 \cos(\omega t) \) is the driving force with amplitude \( F_0 \) and angular frequency \( \omega \). The steady-state particular solution is:
    \begin{equation}
        x(t) = A(\omega)\cos(\omega t - \delta) 
    \end{equation}
    To derive $A(\omega)$ and $\delta$, use the equation of motion:
    \begin{align*}
        A(\omega)(-\omega^2 + \omega_0^2) \cos(\delta) = \omega_0^2 a
    \end{align*}
    where \( a = \frac{F_0}{k} \) is the static displacement of the mass when the driving force is constant. We also have:
    \begin{align*}
        A(\omega)(-\omega^2 + \omega_0^2) \sin(\delta) = 0
    \end{align*}
    Now, consider the case when $\delta = 0$, we have:
    \begin{equation}
        A(\omega) = \frac{\omega_0^2 a}{\omega_0^2 - \omega^2}
    \end{equation}
    and the case where \( \delta = \pi \), we have:
    \begin{equation}
        A(\omega) = -\frac{\omega_0^2 a}{\omega_0^2 - \omega^2}
    \end{equation}
    Observe, when $\omega \approx \omega_0$, the amplitude \( A(\omega) \) becomes very large, indicating resonance. At resonance, the system oscillates with maximum amplitude, which can lead to significant energy transfer from the driving force to the oscillator.
\end{definition}

\subsection{Damped Forced Oscillations}
We add damping to the driven case, derive the frequency response amplitude and phase, and study bandwidth and Q.
\begin{definition}[Damped Forced Oscillations]
    Consider a damped harmonic oscillator subjected to an external driving force. It is described by the following differential equation:
    \begin{equation}
        m\ddot{x} + b\dot{x} + kx = F_0 \cos(\omega t)
    \end{equation}
    where \( F_0 \) is the amplitude of the driving force, \( \omega \) is the angular frequency of the driving force, \( b \) is the damping coefficient, \( m \) is the mass, and \( k \) is the spring constant. The general solution to this non-homogeneous differential equation is given by:
    \begin{equation}
        x(t) = A(\omega)\cos(\omega t - \delta)
    \end{equation}
    and the amplitude \( A(\omega) \) is given by:
    \begin{equation}
        A(\omega) = \frac{\omega_0^2 a}{\sqrt{\omega^2 \gamma^2 + (\omega^2 - \omega_0^2)^2}}
    \end{equation}
    We can derive the following 4 cases:
    \begin{enumerate}
        \item \textbf{No Damping} (\( \gamma = 0 \)): In this case, the amplitude \( A(\omega) \) simplifies to the undamped case we discussed earlier.
        \item \textbf{Low-Frequency Limit} (\( \omega \ll \omega_0 \)): In this limit, the amplitude \( A(\omega) \) approaches the static displacement \( a = \frac{F_0}{k} \). This means that at very low frequencies, the system behaves like a static spring, and the mass is displaced by an amount proportional to the applied force.
        \item \textbf{High-Frequency Limit} (\( \omega \gg \omega_0 \)): In this limit, the amplitude \( A(\omega) \) decreases with increasing frequency, following the relation \( A(\omega) \approx \frac{\omega_0^2 a}{\omega^2} \). This indicates that at very high frequencies, the mass cannot respond quickly enough to the rapidly oscillating driving force, resulting in a smaller amplitude of oscillation.
        \item \textbf{Resonance} (\( \omega \approx \omega_0 \)): At resonance, the amplitude \( A(\omega) \) reaches its maximum value but it does not become infinite due to the presence of damping: 
        \begin{equation}
            A_{\text{max}} = \frac{\omega_0^2 a}{\omega_0 \gamma} = \frac{\omega_0 a}{\gamma} = Q a
        \end{equation}
    \end{enumerate}
    We also have the phase shift \( \delta \) given by:
    \begin{equation}
        \tan(\delta) = \frac{\omega \gamma}{\omega_0^2 - \omega^2}
    \end{equation}
    The phase shift \( \delta \) indicates how much the oscillation of the mass lags behind the driving force. The behavior of \( \delta \) can be summarized as follows:
    \begin{itemize}
        \item At low frequencies (\( \omega \ll \omega_0 \)), \( \delta \) approaches 0, meaning the mass oscillates in phase with the driving force.
        \item At resonance (\( \omega = \omega_0 \)), \( \delta = \frac{\pi}{2} \), indicating that the mass oscillates a quarter cycle behind the driving force.
        \item At high frequencies (\( \omega \gg \omega_0 \)), \( \delta \) approaches \( \pi \), meaning the mass oscillates out of phase with the driving force.
    \end{itemize}
\end{definition}

\begin{definition}[Power absorbed during forced oscillations] 
    We can calculate the velocity of the oscillator:
    \begin{equation}
        \dot{x}(t) = -\omega A(\omega) \sin(\omega t - \delta) = -v_0 \sin(\omega t - \delta)
    \end{equation}
    where \( v_0 = \omega A(\omega) \) is the maximum speed of the oscillator. Energy is lost at the following rate:
    \begin{equation}
        P(t) = b v(t)^2 
    \end{equation}
    Substituteing the expression for \( v(t) \) into the power equation gives:
    $$
        P(t) = b v_0^2 \sin^2(\omega t - \delta)
    $$
    The average power \( \langle P \rangle \) over one complete cycle of the driving force is given by:
   \begin{equation}
        \bar{P}(\omega) = \frac{b v_0^2}{2} = \frac{\omega^2 F_0^2 \gamma}{2m \left[(\omega^2 - \omega_0^2)^2 + \omega^2 \gamma^2\right]}
   \end{equation}
    If the driving frequency \( \omega \) is close to \( \omega_0 \), use $\omega^2 - \omega_0^2 \approx 2\omega_0\Delta \omega$ with $\Delta \omega = \omega - \omega_0$ to get the Lorentzian form:
    \begin{equation}
        \bar{P}(\omega) \approx \frac{\omega_0 F_0^2}{2m \gamma \Big[ 1 + \big(\tfrac{2\Delta \omega}{\gamma}\big)^2 \Big]}
    \end{equation}
    The maximum average power occurs at resonance:
    \begin{equation}
        \bar{P}_{\text{max}} = \frac{\omega_0 F_0^2}{2m \gamma}
    \end{equation}
\end{definition}

\paragraph{Full Width Half Height} The graph of \( \bar{P}(\omega) \) versus \( \omega \) has a peak at \( \omega = \omega_0 \) with a maximum value of \( \bar{P}_{\text{max}} \). The full width at half maximum (FWHM) is the width of the peak at half of its maximum height. The FWHM is given by:
\begin{equation}
    \omega_\text{FWHM} = 2\Delta \omega = 2\gamma = \frac{2\omega_0}{Q}
\end{equation}
To illustrate the graph, we can plot \( \bar{P}(\omega) \) versus \( \omega \) for different values of the quality factor \( Q \):
\begin{figure}[h!]
    \centering
    \begin{tikzpicture}
        \begin{axis}[
            xlabel={Driving Frequency \( \omega \)},
            ylabel={Average Power \( \bar{P}(\omega) \)},
            title={Average Power vs Driving Frequency for Different Quality Factors, Power normalized to \( \bar{P}_{\text{max}}  = 1\)},
            legend pos=north east,
            grid=major,
            width=0.8\textwidth,
            height=0.5\textwidth,
            domain=0:3,
            range=0:1.2,
            samples=200,
        ]
        % Plot for Q = 1
        \addplot[black] {1/(1 + (2*(x-1))^2)};
        \addlegendentry{\( Q = 1 \)}

        % Plot for Q = 5
        \addplot[black, thick, dashed] {1/(1 + (2*(x-1)/0.2)^2)};
        \addlegendentry{\( Q = 5 \)}

        % Plot for Q = 10
        \addplot[black, thick] {1/(1 + (2*(x-1)/0.1)^2)};
        \addlegendentry{\( Q = 10 \)}
        % add the lenght of the fwhm at P = 0.5
        \draw[dashed] (axis cs:0.9,0) -- (axis cs:0.9,0.5);
        \draw[dashed] (axis cs:1.1,0) -- (axis cs:1.1,0.5);
        \draw[<->] (axis cs:0.9,0.5) -- (axis cs:1.1,0.5) node[midway, above] {\( \Delta \omega = \frac{\omega_0}{Q} \)};

        % vertcal line dotted at omega = 1
        \draw[dotted, red, thick] (axis cs:1,0) -- (axis cs:1,1);
        % label the peak
        \node at (axis cs:1,1.05) {\( \bar{P}_{\text{max}} \)};
        % label the x axis at omega = 1
        \node at (axis cs:1,-0.05) {\( \omega_0 \)};

        % label the points
        \node at (axis cs:0.7,-0.05) {\( \omega_0 - \frac{\gamma}{2} \)};
        \node at (axis cs:1.3,-0.05) {\( \omega_0 + \frac{\gamma}{2} \)};

        \end{axis}
    \end{tikzpicture}
    \caption{Average Power vs Driving Frequency for Different Quality Factors}
    \label{fig:power_vs_frequency}
\end{figure}

\subsection{Simple Pendulum}
Consider the arc length \( s \) of a pendulum bob, we have:
\begin{equation}
    s = l\theta \quad m\ddot{s} = -mg\sin(\theta)
\end{equation}

For small angles, we can approximate \( \sin(\theta) \approx \theta \), leading to:
\begin{equation}
    \ddot{\theta} + \frac{g}{l}\theta = 0
\end{equation}
And the ODE has solution:
\begin{equation}
    \theta(t) = \theta_0 \cos(\omega t + \phi)
\end{equation}

\paragraph{Energy} For small angle $\theta$, the total mechanical energy \( E \) of the pendulum is given by:
\begin{equation}
    E = K + U = \frac{1}{2}ml^2\dot{\theta}^2 + mgl(1 - \cos(\theta)) \approx \frac{1}{2}ml^2\dot{\theta}^2 + \frac{1}{2}mgl\theta^2
\end{equation}
This shows that the pendulum behaves like a simple harmonic oscillator with an effective spring constant \( k = \frac{mg}{l} \).

\paragraph{The physical pendulum} For a rigid body swinging about a pivot point, the equation of motion is given by:
\begin{equation}
    I\ddot{\theta} = \tau 
\end{equation}
For simple rod of length \( L \) and mass \( m \) pivoted at one end, the moment of inertia about the pivot point is \( I = \frac{1}{3}mL^2 \). The torque due to gravity when the rod is displaced by an angle \( \theta \) from the vertical is \( \tau = -mg\frac{L}{2}\sin(\theta) \). For small angles, we can approximate \( \sin(\theta) \approx \theta \), leading to:
\begin{equation}
    \ddot{\theta} + \frac{3g}{2L}\theta = 0
\end{equation}
The angular frequency of oscillation for the physical pendulum is:
\begin{equation}
    \omega = \sqrt{\frac{3g}{2L}}
\end{equation}

\paragraph{The LC Circuit} An LC circuit consists of an inductor \( L \) and a capacitor \( C \) connected in series. The charge \( q(t) \) on the capacitor satisfies the differential equation:
\begin{equation}
    L\frac{d^2q}{dt^2} + \frac{1}{C}q = 0
\end{equation}
This is analogous to the equation of motion for a simple harmonic oscillator, with the angular frequency given by:
\begin{equation}
    \omega = \frac{1}{\sqrt{LC}}
\end{equation}

\paragraph{The RLC Circuit} An RLC circuit consists of a resistor \( R \), inductor \( L \), and capacitor \( C \) connected in series. The charge \( q(t) \) on the capacitor satisfies the differential equation:
\begin{equation}
    L\frac{d^2q}{dt^2} + R\frac{dq}{dt} + \frac{1}{C}q = 0
\end{equation}
This is analogous to the resistor behaving as a damped harmonic oscillator, with the angular frequency given by:
\begin{equation}
    \omega = \sqrt{\frac{1}{LC} - \left(\frac{R}{2L}\right)^2}
\end{equation}

\section{The Wave Equation}

\begin{definition}[The Wave Equation]
    The wave equation is a second-order linear partial differential equation that describes the propagation of waves (sound, EM, water). In one dimension, it is:
    \begin{equation}
        \frac{\partial^2 u}{\partial t^2} - c^2 \frac{\partial^2 u}{\partial x^2} = 0
    \end{equation}
    where $u(x,t)$ is the disturbance, $c$ is the wave speed, $x$ is position, and $t$ is time.
\end{definition}

\chapter{Modern Physics}
\end{document}